---
title: "3_R_advance_analysis_visualization"
author: "Roi & Michael"
date: "09/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load(file = "./training_data/trainin_eflalo_tacsat.RData")

```



## Advanced R using Tidyverse: Tibble, dplyr, ggplot2 

### Reading and Writing data to and from R

R can deal with a number of file types, many can be loaded/saved using simple 'save' and 'load' functions from the base package, but there are some file type specific functions 

#### RData files
```{r rdata, message=FALSE , warning= FALSE}
  save(mm, file = "./output/mm.RData") # saving an R object as an RData file so it can be shared or imported into another project
  rm(mm)
  
  load("./output/mm.RData") # load in the RData file that we just saved and removed
```

#### CSV files
```{r csv, message=FALSE , warning= FALSE}
  write.csv(mm, file = "./output/mm.csv") # write to csv format
  mmcsv <- read.csv("./output/mm.csv") # read the csv and name the variable
```

#### Excel files
```{r xlsx, message=FALSE , warning= FALSE}
install.packages("xlsx")
library(xlsx)

write.xlsx(mm, file = "./output/mm.xlsx") # writes an object to xlsx format
mmxl <- read.xlsx("./output/mm.xlsx", sheetName = "Sheet1") # reads in data from an xlsx format
```

### Tidyverse package and attached packages 

library(tidyverse) ## loads ggplot2 a, tibble , tidyr and dplyr


#### a.Tibbles 

Throughout this book we work with “tibbles” instead of R’s traditional data.frame. Tibbles are data frames, but they tweak some older behaviours to make life a little easier. R is an old language, and some things that were useful 10 or 20 years ago now get in your way. It’s difficult to change base R without breaking existing code, so most innovation occurs in packages. Here we will describe the tibble package, which provides opinionated data frames that make working in the tidyverse a little easier. In most places, I’ll use the term tibble and data frame interchangeably; when I want to draw particular attention to R’s built-in data frame, I’ll call them data.frames.

[https://r4ds.had.co.nz/tibbles.html](url)

##### **Example**

```{r tibbles, message=FALSE , warning= FALSE}

   library(tidyverse)

  # Create a tibble from scratch 

  tibble ( 
    x = 1:5, 
    y = 1, 
    z = x^2+y 
    )

  # Convert a data.frame into a tibble

   fish_d <- fish_data%>%as_tibble()
   
  # Print more than 10 tibbles rows
   
   fish_d%>%print()   ## Default print 10 rows
   fish_d%>%print(n = 15, width = Inf)
   
  
  # Pull a data column from a tibble and convert it into a vector
   
   fish_d%>%select(SI_SP)%>%pull()%>%head(100)
   
  # Rename a tibble column 
   
   fish_d%>%rename( vessel_speed = SI_SP, fishing_time = INTV)
   
   

```


# b.Dplyr and analysis datasets   ------

Tibbles is the tidyverse standard table format , all results form tidyverse package will be in a tibble format. 
Dplyr is the tidyverse package used to modify , calcualte and analyse the data in a tibble. 

Dplyr sintax is very similar to SQL , so the methods and uses of Dplyr functions are very similar to SQL language. 
Let's explore some of them.


```{r dplyr, message= FALSE}

# Tibble manipulation

fish_d%>%select( LE_GEAR, LE_MET , INTV)   # select 3 fields 
fish_d%>%select( -LE_MET )   ### select all minus one of the columns


fish_d%>%mutate(sum_fields = SI_SP + INTV) # Create a new field as result of calculation


### Average calculation by groups

fish_d%>%group_by(LE_GEAR)%>%summarise(mean = mean(SI_SP), max = max(SI_SP), min = min(SI_SP))



```



# c.ggplot2 -------

ggplot2 is an extremely useful package when it comes to plotting data. for example, the following code creates a column graph depicting the total weight caught by each egar type, in the fish_data dataset.

```{r ggplot2, message= FALSE}
unique(fish_data$LE_GEAR) # Lists the gear tyes in the dataset

ggplot(data = fish_data) +
  geom_col(mapping = aes(x = LE_GEAR, y = LE_KG_TOT))
```

Now the value.

```{r ggplot2, message= FALSE}
ggplot(data = fish_data) +
  geom_col(mapping = aes(x = LE_GEAR, y = LE_EURO_TOT))
```

Use of par enables you to organise the outputs, so they can be viewed side by side.

```{r ggplot2, message= FALSE}
par(mfrow=c(1,2))

ggplot(data = fish_data) +
  geom_col(mapping = aes(x = LE_GEAR, y = LE_KG_TOT))

ggplot(data = fish_data) +
  geom_col(mapping = aes(x = LE_GEAR, y = LE_KG_TOT))
```

Colours can be changed with a simple color command.

```{r ggplot2, message= FALSE}
par(mfrow=c(1,2))

ggplot(data = fish_data) +
  geom_col(mapping = aes(x = LE_GEAR, y = LE_KG_TOT), color = "blue")

ggplot(data = fish_data) +
  geom_col(mapping = aes(x = LE_GEAR, y = LE_KG_TOT), colour = "orange")
```







