---
title: "eflalo_tacsat_analysis"
author: "Roi & Mike"
date: "2023-09-15"
output: html_document
---

```{r setup, include=TRUE, eval=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(vmstools)
library(dplyr)
library(ggplot2)
library(sf)
library(tidyr)
```

Set the working directory to the data folder, if it is not already.

```{r set wd, include=TRUE, eval=FALSE}
setwd('./../../data')
```

Load the data made through the previous scripts.

```{r load data, include=TRUE, eval=FALSE}
  load(file = '.\\workflow_outputs\\eflalo_fs_qc.RData' )
  load(file = '.\\workflow_outputs\\tacsat_fs_qc.RData' )
```

### 1. Merge the Tacsat and Eflalo data together

Explore the data

```{r explore, include=TRUE, eval=FALSE}
  str(tacsat_fs)
  str(eflalo_fs)
 
  dim(eflalo_fs)
```

Convert EFLALO from LARGE format into WIDE format. This complies with the format required by VMSTOOL 'splitamongpings' function. Depending on the format selected, we have to choose Option 1 or Option 2 in step '2.3 Dispatch landings/catches among VMS pings'. As a general approach, if we want to analyse all species together choose long format if we want to analyse a number of species choose wide format.

```{r pivot eflalo, include=TRUE, eval=FALSE}
  eflalo_fs = eflalo_fs%>%
    pivot_wider(id_cols = c( FT_REF, FT_DCOU, FT_DHAR, FT_DDAT, FT_DTIME, FT_DDATIM,
                             FT_LCOU, FT_LHAR, FT_LDAT, FT_LTIME, FT_LDATIM, VE_REF,
                             VE_FLT, VE_COU, VE_LEN, VE_KW, VE_TON, FT_YEAR,
                             LE_ID, LE_CDAT, LE_STIME, LE_ETIME, LE_SLAT, LE_GEAR,
                             LE_MSZ, LE_RECT, LE_DIV, LE_MET, EFLALO_FT_FT_REF, Year,
                             Month, VE_LEN_CAT, SOURCE, FLEET_SEG ),
                names_from = c(LE_SPE), values_from = c(LE_KG, LE_VALUE))
 
 
  dim(eflalo_fs)
```

Retrieve a selection of columns from eflalo_fs to be joined to tacsat.

```{r eflalo_sel, include=TRUE, eval=FALSE}
   eflalo_sel =  eflalo_fs %>% 
                  select(FT_REF, LE_CDAT, LE_GEAR,LE_MSZ, LE_RECT, LE_MET, VE_LEN, VE_KW, VE_COU) %>%
                  distinct()
```

Join the data and check all gears transferred correctly. The number of NA values in the LE_GEAR column must be 0.

```{r join tacsat eflalo_sel, include=TRUE, eval=FALSE}
    tacsatp =  tacsat_fs%>%  left_join( eflalo_sel  , by = c("SI_FT" = "FT_REF", "SI_DATE" = "LE_CDAT"))
  
    tacsatp%>%filter (is.na(LE_GEAR)) ## must be 0 , that means all iVMS records have an associated EFLALO records 
```

Retrieve the first trip from this list after rearranging the data. 

```{r , include=TRUE, eval=FALSE}
  ft =    eflalo_sel %>% arrange(FT_REF , LE_CDAT) %>% 
          select (FT_REF, LE_GEAR) %>% rename(LE_GEAR2 = LE_GEAR) %>%
          group_by(FT_REF)%>%slice(1)
```

Assign the main gear of each trip to the VMS locations that do not already have a gear assigned.

```{r , include=TRUE, eval=FALSE}
  tacsatp =  tacsatp %>% left_join ( ft , by = c( "SI_FT" = "FT_REF"))%>%
    mutate ( LE_GEAR = ifelse( is.na (LE_GEAR), LE_GEAR2 , LE_GEAR  ))%>%
    select (-LE_GEAR2)
```

Check there are no NA records in the LE_GEAR column

```{r check gear NAs, include=TRUE, eval=FALSE}
tacsatp %>% filter( is.na( LE_GEAR ) )
```

Input the information for vessel characteristics for locations that do not have an associated log event.

```{r vessel characteristics, include=TRUE, eval=FALSE}
  ve_c = eflalo_sel%>%distinct(FT_REF, VE_LEN, VE_KW, VE_COU)%>%rename(VE_LEN2 = VE_LEN, VE_KW2 = VE_KW, VE_COU2 = VE_COU )
  
  tacsatp =  tacsatp %>% left_join ( ve_c , by = c( "SI_FT" = "FT_REF"))%>%
    mutate ( VE_LEN = ifelse(is.na (VE_LEN), VE_LEN2 , VE_LEN  ), 
             VE_KW = ifelse( is.na (VE_KW), VE_KW2 , VE_KW  ) , 
             VE_COU = ifelse(is.na (VE_COU), VE_COU2 , VE_COU  ) )%>%
    select (-c(VE_LEN2, VE_KW2, VE_COU2))
  
  
  tacsatp %>% filter(is.na(VE_LEN))
  
  tacsatp = tacsatp %>% as.data.frame() # convert to data frame for further analysis
```

### 2. Define Activity

#### 2.1 Calculate interval between VMS points. This information can be used to calculate effort.

```{r define activity, include=TRUE, eval=FALSE}
  minInterval = 1 / 60 ## 1 minute converted in hours (0.01666667 hours)
  
  tacsatp = tacsatp%>%ungroup()%>%
    group_by(VE_REF, SI_FT)%>% arrange (VE_REF, SI_DATIM)%>%
    mutate (INTV =  difftime(SI_DATIM , lag(SI_DATIM), units = "hours" )  )%>%  ## Calculate the difference between a iVMS loction time stamp and previous location to calcualte a fishign effort in a given location
    mutate(interval_mean = mean(INTV , na.rm = T))%>%                           ## Calculate the mean to replace the NA's interval when a VMS location is the 1st of a trip and cannot calculate with a prev. iVMS location
    mutate(INTV = ifelse( is.na(INTV), interval_mean, INTV ))%>%                ## Convert the NA's into a effort represented by the mean of that vessel durign given trip
    select(- interval_mean) %>%ungroup()
  

  # Review the  INTV that are too high e.g. 5x the regular interval rate by trip 
  
  tacsatp %>% group_by(SI_FT) %>% mutate  (   mean_intv_trip = mean ( INTV  )) %>%
    filter ( INTV  > 5 * mean_intv_trip)
```

#### 2.2 Define speed thresholds for gears

Plot a histogram. Histograms are a useful way of visualising the data so that abnormalities or patterns can be observed.

```{r speed histogram, include=TRUE, eval=FALSE}
  ggplot(data = tacsatp, aes(SI_SP)) +
    geom_histogram(
      breaks = seq(0, 10, by =0.4), col = 1) +
    facet_wrap( ~ LE_GEAR, ncol = 4, scales = "free_y") +
    labs(x = "Speed (knots)", y = "Frequency") +
    theme(
      axis.text.y = element_text(colour = "black"),
      axis.text.x = element_text(colour = "black"),
      axis.title.y = element_text(size = 14),
      axis.title.x = element_text(size = 14),
      panel.background = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.line = element_line(colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA)
    )
```

Create a speed threshold for what is to be considered realistic for fishing activity. This can be chosen based on the histogram, which will show outliers which are unrealistic, this threshold can eliminate those.

```{r speed threshold, include=TRUE, eval=FALSE}
  speedarr = as.data.frame( cbind( LE_GEAR = sort(unique(tacsatp$LE_GEAR)), min = NA, max = NA), stringsAsFactors = FALSE)
  if (!'MIS' %in%  unique (speedarr$LE_GEAR ) ) {  speedarr =  speedarr %>% rbind(c('MIS', NA, NA))  } 
```

Fill the speed array with min and max values as a general rule of thumb or with specific values per gear.

```{r speedarr, include=TRUE, eval=FALSE}
  speedarr$min = rep(1, nrow(speedarr)) # Filled with rule of thumb 1 knot as minimum fishing speed
  speedarr$max = rep(6, nrow(speedarr)) # Filled with rule of thumb 6 knot as maximum fishing speed
  
  ## Change only one gear
  
  speedarr = speedarr %>% mutate( min = if_else(LE_GEAR == 'GN', 0, min ), 
                                  max = if_else(LE_GEAR == 'GN', 1.5, max ))
```

Analyse activity for common gears. The speed array can be used for other gears.

```{r autodetect gears, include=TRUE, eval=FALSE}
  autoDetectionGears = c('OTB',  'DRB')
 
  subTacsat    = subset(tacsatp, LE_GEAR %in% autoDetectionGears)%>%as.data.frame()
  nonsubTacsat = subset(tacsatp, !LE_GEAR %in% autoDetectionGears)%>%as.data.frame()
```

Create the storeScheme object where the speed profiles will be stored.

```{r storeScheme, include=TRUE, eval=FALSE}
    storeScheme =  expand.grid(  years = 2022, months = 0,  weeks = 0,  analyse.by = unique(subTacsat[, "LE_GEAR"]) )
    
    storeScheme$peaks <- NA
    storeScheme$means <- NA
    storeScheme$fixPeaks <- FALSE
    storeScheme$sigma0 <- 0.911
```

Fill the storeScheme with values based on analysis of the graphs produced.

```{r fill storeScheme, include=TRUE, eval=FALSE}
    dat = subTacsat %>% filter ( LE_GEAR == 'OTB' )   
    
    datmr = dat %>%mutate (SI_SP = -1 * SI_SP )
    datmr = dat %>% bind_rows(datmr)
    
    xrange = pmin(20, range(datmr$SI_SP), na.rm = TRUE)
    datmr$SI_SP[which(abs(datmr$SI_SP) > 20)] =  NA
    hist(datmr$SI_SP, breaks = seq(xrange[1]- 3, xrange[2] + 3 , 0.5),  main = paste('OTB'), xaxt = "n")
    axis(1, at = seq(xrange[1]- 3, xrange[2] + 3, 1))
```

Define the mean values of the peaks and the number of peaks when they are different from 5.

```{r define means, include=TRUE, eval=FALSE}
    storeScheme$means[which(storeScheme$analyse.by == "DRB")] <- c("-9 -4.5 0 4.5 9")
    storeScheme$peaks[which(storeScheme$analyse.by == "DRB")] <- 5
    storeScheme$means[which(storeScheme$analyse.by == "OTB")] <- c("-9 -4.5 0 4.5 9")
    storeScheme$peaks[which(is.na(storeScheme$peaks) == TRUE)] <- 5
    
    subTacsat%>%filter(LE_GEAR == 'OTB')
```

Generate graph for the above data.

```{r acTa, include=TRUE, eval=FALSE}
  acTa = activityTacsat( subTacsat ,units = "year",  analyse.by = "LE_GEAR",storeScheme = storeScheme, plot = TRUE, level = "all")
  subTacsat$SI_STATE = acTa
  subTacsat$ID <- 1:nrow(subTacsat)
  
  subTacsat%>%group_by(LE_GEAR, SI_STATE)%>% summarise(f_min_s = min (SI_SP ), f_max_s = max (SI_SP ))
  ggplot ( subTacsat%>%filter ( LE_GEAR== 'OTB') , aes( x= SI_SP, fill = SI_STATE)  ) + geom_bar ( ) +theme_minimal()
```

Check the results of the above, if they are not satisfactory, the analysis can be run again with fixed peaks.

```{r autodetection, include=TRUE, eval=FALSE}
  for (iGear in autoDetectionGears) {
    
    subDat <- subset(subTacsat,LE_GEAR == iGear)
    minS = min(subDat$SI_SP[which(subDat$SI_STATE == "s")],  na.rm = TRUE) ## minimum steaming speed
    minF = min(subDat$SI_SP[which(subDat$SI_STATE == "f")],  na.rm = TRUE) ## minimum fishing speed
    
    if(minS < minF) {
      storeScheme$fixPeaks[which(storeScheme$analyse.by == iGear)] = TRUE
      subacTa = activityTacsat(
          subDat, units = "year", analyse.by = "LE_GEAR", storeScheme, plot = FALSE,level = "all"
        )
      subTacsat$SI_STATE[subDat$ID] = subacTa
    }
  }
  
  subTacsat =  subTacsat%>%select(-ID)
  
  subTacsat%>%group_by(LE_GEAR, SI_STATE)%>% summarise(f_min_s = min (SI_SP ), f_max_s = max (SI_SP ))
  
  ggplot ( subTacsat%>%filter ( LE_GEAR== 'OTB') , aes( x= SI_SP, fill = SI_STATE)  ) + geom_bar ( ) +theme_minimal()
  ggplot ( subTacsat%>%filter ( LE_GEAR== 'DRB') , aes( x= SI_SP, fill = SI_STATE)  ) + geom_bar (width = 0.1 ) + theme_minimal()
```

Assign a speed rule classification for the gears which were visually inspected.

```{r speed rule classification, include=TRUE, eval=FALSE}
  ## This will apply the values in speed array data frame created at the beginning of the process
  
  metiers = unique(nonsubTacsat$LE_GEAR)
  nonsubTacsat$SI_STATE <- NA
  
  for (mm in metiers) {
    
    sp_gear =  speedarr %>%filter(LE_GEAR == mm) 
    nonsubTacsat = nonsubTacsat %>% 
                    mutate ( SI_STATE = if_else  ( LE_GEAR == mm & SI_SP >= sp_gear$min & SI_SP <= sp_gear$max, 'f','s' ))
      
   
  }
```

If there are still NAs, MIS must be assigned to rows which are within the fishing threshold.

```{r replace NAs with MIS, include=TRUE, eval=FALSE}
  sp_gear =  speedarr %>%filter(LE_GEAR == 'MIS')
  
  nonsubTacsat =  nonsubTacsat %>% 
                  mutate ( SI_STATE = if_else  ( is.na(LE_GEAR)  & SI_SP >= sp_gear$min & SI_SP <= sp_gear$max, 'f','s' ))

  
  
  nonsubTacsat =  nonsubTacsat%>% mutate (SI_STATE = if_else ( is.na(SI_STATE), 's', SI_STATE)  )  
  
  
  # Combine the two dataset together again into TACSATP =============== 
  
  
  tacsatp = subTacsat %>% bind_rows( nonsubTacsat )%>% arrange(VE_REF, SI_DATIM)   
```

Set fishing sequences with hauling in the middle of them to 'f'.

```{r hauling sequences, include=TRUE, eval=FALSE}
  idx =   which(
            tacsatp$SI_STATE[2:(nrow(tacsatp) - 1)] == "h" &
              tacsatp$SI_STATE[1:(nrow(tacsatp) - 2)] == "f" &
              tacsatp$SI_STATE[3:(nrow(tacsatp))    ] == "f" &
              tacsatp$VE_REF[2:(nrow(tacsatp) - 1)] == tacsatp$VE_REF[1:(nrow(tacsatp) - 2)] &
              tacsatp$VE_REF[2:(nrow(tacsatp) - 1)] == tacsatp$VE_REF[3:(nrow(tacsatp))]
           ) + 1
  
  tacsatp$SI_STATE[idx] <- "f"
```

Save.

```{r save tacsat activity, include=TRUE, eval=FALSE}
  save( tacsatp, file = paste0(".\\workflow_outputs\\tacsatActivity", year, ".RData")) 
  
  message("Defining activity completed")
```

Depending on the format of the eflalo data, wide or long, the analysis will need to be different, so select the type being used here. Also, the results from the data and how their weights per species are summarised can be changed.

```{r wide/long, include=TRUE, eval=FALSE}
  eflalo_format = 'wide' ## default format following toolbox workflow 
  species_analysis_type = 'all_species_sum' # options: ( all_species_sum,  selected_species_sum, selected_species_separated )
  
  if ( eflalo_format == 'wide' ) { 
  
  ## For all the species together 
  
  
  ## Option 1: Names in wide format . Each species catch value has its own column 
 
  idxkg  =  grep("KG", colnames(eflalo_fs ) )  
  idxval =  grep("VALUE", colnames(eflalo_fs )  ) 
  
  
  
  ## Option 1.1 : all landings for all species sum together 
  if (species_analysis_type == 'all_species_sum') {
  
  eflalo_fs_tot = eflalo_fs %>% mutate(LE_KG_TOT = select(.,idxkg) %>% rowSums(., na.rm = TRUE) ) %>% select ( -c ( idxkg, idxval) ) %>% mutate(LE_EURO_TOT = NA ) %>% as.data.frame()       
  
  } else if (species_analysis_type == 'selected_species_sum') {
    
  ## Option 1.2 : by selected species totals
  eflalo_fs_tot = eflalo_fs %>%   mutate(LE_KG_TOT = select(.,contains(c('KG_SOL', 'KG_MAC')) ) %>% rowSums(., na.rm = TRUE) ) %>% select ( -c ( idxkg, idxval) ) %>% as.data.frame()       
  
  } else if (species_analysis_type == 'selected_species_separated') {
  
  ## Option 1.3 : Analysis by species separately 
  eflalo_fs_tot = eflalo_fs %>% select(., -idxkg , -idxval, contains(c('SOL', 'MAC') ) ) %>% filter_at ( vars (contains( c('SOL', 'MAC') ) ) , any_vars(!is.na(.)) ) %>%  as.data.frame()
  
  }
  
  dim( eflalo_fs_tot)
  
  }  else if   (eflalo_format == 'long'  ) { 
    
      ## Option 2: Species names in long format. All  species in LE_SPE and  catch value in LE_KG columns
  
  
  captures_total = eflalo_fs %>% group_by(FT_REF, LE_ID) %>% summarise(LE_KG_TOT = sum( LE_KG), LE_EURO_TOT = 0 ) # LE_KG_TOT = sum( LE_KG) if VALUE data is available
 

  eflalo_fs_tot = eflalo_fs %>% select(-c (LE_SPE, LE_KG, LE_VALUE)) %>% distinct() %>% inner_join(captures_total , by = c("FT_REF", "LE_ID" ))
  
  } 
```

Save the newly processed eflalo data.

```{r save, include=TRUE, eval=FALSE}
  save( eflalo_fs_tot, file = paste0(".\\workflow_outputs\\eflaloTotals", year, ".RData")) 
```

Convert the SI_FT column to FT_REF to meet the requirements of the functions. Also, investigate which trips are not merged. Finally, reassign fishing/non-fishing from 'f' and 's' to 1 and 0.

```{r state to 1 or 0, include=TRUE, eval=FALSE}
  tacsatp = tacsatp %>% mutate( FT_REF = SI_FT )
  
  eflaloM =  subset(eflalo_fs_tot,FT_REF %in% unique(tacsatp$FT_REF))
  eflaloNM = subset(eflalo_fs_tot,!FT_REF %in% unique(tacsatp$FT_REF))
  
  dim(eflaloM)
  dim(eflaloNM)
  
  
  tacsatp$SI_STATE[which(tacsatp$SI_STATE != "f")] = 0
  tacsatp$SI_STATE[which(tacsatp$SI_STATE == "f")] = 1
```

Create tacsatEflalo. Take the records which are considered to be fishing.

```{r create tacsatEflalo, include=TRUE, eval=FALSE}
  tacsatEflalo = tacsatp[tacsatp$SI_STATE == 1,] 
  
  tacsatEflalo = tacsatEflalo %>% filter ( FT_REF %in%  ( eflaloM%>%distinct(FT_REF)%>%pull())   )
```

Replace NA values in weight and vlue columns with 0s.

```{r weight/value 0s, include=TRUE, eval=FALSE}
  eflaloM = eflaloM %>% mutate ( LE_VALUE_TOT = 0 )
  eflaloM = eflaloM %>% mutate ( LE_EURO_TOT = LE_VALUE_TOT   )
```

Run the split among pings function and save the result.

```{r split among pings, include=TRUE, eval=FALSE}
  ?splitAmongPings
 
    tacsatEflalo =
      splitAmongPings(
        tacsat = tacsatEflalo,
        eflalo = eflaloM,
        variable = "kgs", # "all",
        level = "day",
        conserve = FALSE
      )
 
    tacsatEflalo%>%filter(!is.na( LE_KG_TOT  ) ) 
    
    eflaloM%>%summarise(TOT = sum(LE_KG_TOT))
    tacsatEflalo%>%summarise(TOT = sum(LE_KG_TOT))
    
    
  
  save(
    tacsatEflalo,
    file = paste0("workflow_outputs\\tacsatEflalo", year, ".RData")
  )
  
  print("Dispatching landings completed")
```

Investigate a specific trip in all 3 datasets.

```{r , include=TRUE, eval=FALSE}
  tacsatp%>% filter(SI_STATE == 1 & SI_FT == 610736051)%>% dim()
  eflaloM %>% filter(FT_REF == 610736051)
  tacsatEflalo%>% filter(FT_REF == 610736051) %>% summarise(tot = sum ( LE_KG_TOT))
```

Investigate the data at vessel level.

```{r , include=TRUE, eval=FALSE}
  eflaloM %>% filter(VE_REF == 'A10207')%>% summarise(tot = sum ( LE_KG_TOT))
  tacsatEflalo%>% filter(VE_REF == 'A10207') %>% summarise(tot = sum ( LE_KG_TOT))
```

Assign c-square, year, month, quarter, area and create tacsatEflalo_output for the year of study. Then, save.

```{r , include=TRUE, eval=FALSE}
  tacsatEflalo$Csquare   =  CSquare(tacsatEflalo$SI_LONG, tacsatEflalo$SI_LATI, degrees = 0.05)
  tacsatEflalo$Year      =  year(tacsatEflalo$SI_DATIM)
  tacsatEflalo$Month     =  month(tacsatEflalo$SI_DATIM)
  tacsatEflalo$kwHour    =  as.numeric(tacsatEflalo$VE_KW) * tacsatEflalo$INTV  
  tacsatEflalo$cpue      =  tacsatEflalo$LE_KG_TOT / tacsatEflalo$INTV   
  tacsatEflalo$INTV      =  tacsatEflalo$INTV 
  
  
  
  save(tacsatEflalo, file = paste0("workflow_outputs\\tacsatEflalo_output_", year, ".RData"))
```

Assign  year, month, quarter, area and create eflalo_output for the year of study. Then, save.

```{r , include=TRUE, eval=FALSE}
  eflalo_output$Year      = year(eflalo_output$FT_LDATIM)
  eflalo_output$Month     = month(eflalo_output$FT_LDATIM)
  eflalo_output$INTV      = 1 # 1 day
  
  eflalo_output = eflalo_output %>% group_by(VE_REF, LE_CDAT) %>% mutate ( nr = row_number() ) %>% 
                  mutate ( nr = max(nr) ) %>% 
                  mutate ( INTV = INTV / nr)   %>% 
                  arrange ( VE_REF , LE_CDAT)  %>% as.data.frame()
  
   
  
  eflalo_output$kwDays =  as.numeric(eflalo_output$VE_KW) * eflalo_output$INTV
  eflalo_output$tripInTacsat = ifelse(eflalo_output$FT_REF %in% tacsatp$FT_REF, "Y", "N") # Y = Yes and N = No
  
  save(eflalo_output,file = paste0("workflow_outputs\\eflalo_output_", year, ".RData"))
```