---
title: "2_eflalo_tacsat_analysis"
author: "Roi & Mike"
date: "22/03/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lubridate)
library(dplyr)
library(sf)
library(ggplot2)
library(vmstools)
```

## Merge the eflalo and tacsat data together 

### Load in the previously cleaned eflalo and tacsat data

```{r load data, include=FALSE}
load(file = '.\\..\\data\\eflalo_gbw.RData' )
load(file = '.\\..\\data\\tacsat_gbw.RData' )
```

To ensure columns match between datasets, the SI_FT column will be renamed to FT_REF. This uses mutate to create a new column named FT_REF, instead of replacing the SI_FT column. The column name is specified before the = and the contents are specified after.

```{r SI_FT to FT_REF, include=FALSE}
tacsatp = tacsat_gbw %>% mutate(FT_REF = SI_FT) 
  
head(tacsat_gbw) # observe the new column
```

### Assign gear and length to tacsat

This code simply retrieves the information from the eflalo dataset and applies it to new columns created in the tacsat dataset, using the matching FT_REF values. to the left of the arrow, the new column is made and named. To the right, it is saying where the information must come from, followed by the criteria inside the brackets. Using the match function, matching FT_REF values are identified and the information is pulled from eflalo and placed in the new tacsat column.

```{r gear and length, include=FALSE}
  tacsatp$LE_GEAR  <- eflalo$LE_GEAR  [ match(tacsatp$FT_REF, eflalo$FT_REF)]
  tacsatp$LE_MSZ   <- eflalo$LE_MSZ   [ match(tacsatp$FT_REF, eflalo$FT_REF)]
  tacsatp$VE_LEN   <- eflalo$VE_LEN   [ match(tacsatp$FT_REF, eflalo$FT_REF)]
  tacsatp$VE_KW    <- eflalo$VE_KW    [ match(tacsatp$FT_REF, eflalo$FT_REF)]
  tacsatp$LE_RECT  <- eflalo$LE_RECT  [ match(tacsatp$FT_REF, eflalo$FT_REF)]
  tacsatp$LE_MET   <- eflalo$LE_MET   [ match(tacsatp$FT_REF, eflalo$FT_REF)]
  tacsatp$LE_WIDTH <- eflalo$LE_WIDTH [ match(tacsatp$FT_REF, eflalo$FT_REF)]
  tacsatp$VE_FLT   <- eflalo$VE_FLT   [ match(tacsatp$FT_REF, eflalo$FT_REF)]
  tacsatp$LE_CDAT  <- eflalo$LE_CDAT  [ match(tacsatp$FT_REF, eflalo$FT_REF)]
  tacsatp$VE_COU   <- eflalo$VE_COU   [ match(tacsatp$FT_REF, eflalo$FT_REF)]
```

### Define activity

Reset values which are over twice the regular interval rate. Set a threshold of 2, which will be used as a multiplier and calculate each interval multiplied by the threshold value.

```{r interval threshold, include=FALSE}
intvThres = 2
  
tacsatp$INTV[tacsatp$INTV > intvThres] <- 2 * intvThres
```

Remove points with NA values in critical places. This includes numerical columns which will be used for calculations. NA values will often cause the result of any calculations to return as NA, making them useless. By removing these, the calculations can go ahead.

```{r remove critical NAs, include=FALSE}
  head(tacsatp) # check the data first

# remove NAs from these specified columns
  tacsatp%>%filter(
    !is.na(VE_REF)  |
    !is.na(SI_LONG) | 
    !is.na(SI_LATI) |
    !is.na(SI_DATIM)| 
    !is.na(SI_SP)
  )
```

Histograms are a useful way of visualising the data so that abnormalities or patterns can be observed.

```{r plot speed histogram, include=FALSE}
  png(filename = file.path(outPath, paste0("SpeedHistogram_", year, ".png"))) # read in the image using the png function

  # plot the tacsat data using the speed column (SI_SP)
  ggplot(data = tacsatp, aes(SI_SP)) + # selecting the data and the aesthetics (what specifically will be plotted)
    geom_histogram( # choose the type of plot, in this case, histogram
      breaks = seq(0, 20, by =0.4), col = 1) + # specify how the data will be broken up. This scale will be 0-20 in increments of 0.4
    facet_wrap( ~ LE_GEAR, ncol = 4, scales = "free_y") +
    labs(x = "Speed (knots)", y = "Frequency") + # label the axis
    
    # setting the colour options for text and the main elements of the plot
    theme(
      axis.text.y = element_text(colour = "black"),
      axis.text.x = element_text(colour = "black"),
      axis.title.y = element_text(size = 14),
      axis.title.x = element_text(size = 14),
      panel.background = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.line = element_line(colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA)
    )
  dev.off()
```

Create a speed threshold for what is to be considered realistic for fishing activity. This can be chosen based on the histogram, which will show outliers which are unrealistic, this threshold can eliminate those.

```{r speed threshold, include=FALSE}
  speedarr <-
    as.data.frame(
      cbind(
        LE_GEAR = sort(unique(tacsatp$LE_GEAR)), # Retrieves all the unique entries in the LE_GEAR column and sorts them alphabetically
        min = NA,
        max = NA),
      stringsAsFactors = FALSE)

# set the range here, with the minimum and maximum speeds
  speedarr$min <- rep(1, nrow(speedarr)) # It is important to fill out the personally inspected thresholds here!
  speedarr$max <- rep(6, nrow(speedarr))
```



```{r Analyse gears, include=FALSE}

# create two subsets, one where the gear listed in the tacsat data is present in the autoDetectionGears data, and one where it is not
  subTacsat <- subset(tacsatp, LE_GEAR %in% autoDetectionGears)
  nonsubTacsat <- subset(tacsatp, !LE_GEAR %in% autoDetectionGears)
  
  if (visualInspection == TRUE) # set a criteria which must be met for the code to run
  {
    storeScheme <-
      activityTacsatAnalyse(
        subTacsat,
        units = "year",
        analyse.by = "LE_GEAR",
        identify = "means")
  } else # if the criteria was not met for the above code to run, the following code will be executed.
  {
    storeScheme <-
      expand.grid(
        years = year,
        months = 0,
        weeks = 0,
        analyse.by = unique(subTacsat[, "LE_GEAR"])
      )
    storeScheme$peaks <- NA
    storeScheme$means <- NA
    storeScheme$fixPeaks <- FALSE
    storeScheme$sigma0 <- 0.911
    
        
    
    # Fill the storeScheme values based on analyses of the pictures = 
    
    
    # Define mean values of the peaks and the number of peaks when they are different from 5 # 
    
    
    storeScheme$means[which(storeScheme$analyse.by == "TBB")] <- c("-11.5 -6 0 6 11.5")
    storeScheme$means[which(storeScheme$analyse.by == "OTB")] <- c("-9 -3 0 3 9")
    storeScheme$means[which(storeScheme$analyse.by == "OTT")] <- c("-9 -3 0 3 9")
    storeScheme$means[which(storeScheme$analyse.by == "SSC")] <- c("-9 0 9")
    storeScheme$means[which(storeScheme$analyse.by == "PTB")] <- c("-10 -3 0 3 10")
    storeScheme$means[which(storeScheme$analyse.by == "DRB")] <- c("-10 0 10")
    storeScheme$means[which(storeScheme$analyse.by == "HMD")] <- c("-9 0 9")
    storeScheme$peaks[which(storeScheme$analyse.by == "SSC")] <- 3
    storeScheme$peaks[which(storeScheme$analyse.by == "DRB")] <- 3
    storeScheme$peaks[which(storeScheme$analyse.by == "HMD")] <- 3
    storeScheme$peaks[which(is.na(storeScheme$peaks) == TRUE)] <- 5
  }
  
  acTa <-
    activityTacsat(
      subTacsat,
      units = "year",
      analyse.by = "LE_GEAR",
      storeScheme = storeScheme,
      plot = FALSE,
      level = "all")
  subTacsat$SI_STATE <- acTa
  subTacsat$ID <- 1:nrow(subTacsat)
```

Check the results of the previous section. If the results are not satisfactory, rerun the analysis with fixed peaks.

```{r check results of gear analysis, include=FALSE}
  for (iGear in autoDetectionGears) { # open a for loop to iterate through each item in the autoDetectionGears dataset
    subDat <- subset(subTacsat,LE_GEAR == iGear) # create a data subset which takes the current gear type in the loop (iGear)
    minS <-
      min(
        subDat$SI_SP[which(subDat$SI_STATE == "s")], # create a sub dataset which includes only steaming entries 
        na.rm = TRUE)
    minF <-
      min(subDat$SI_SP[which(subDat$SI_STATE == "f")], # do the same for fishing 
          na.rm = TRUE)
    if(minS < minF) { # this part of the function will only work if the criteria is met
      storeScheme$fixPeaks[which(storeScheme$analyse.by == iGear)] <- TRUE
      subacTa <-
        activityTacsat( 
          subDat,
          units = "year",
          analyse.by = "LE_GEAR",
          storeScheme,
          plot = FALSE,
          level = "all"
        )
      subTacsat$SI_STATE[subDat$ID] <- subacTa
    }
  }
  subTacsat <-
    subTacsat[,
              -rev(grep("ID", colnames(subTacsat)))[1]
    ]
```

Assign min or max to the sows of SI_SP and progress with analysis. Entries will be tested to see if they meet set criteria. In this case, the gear type is checked; if it is within the metiers dataset which is created below, and if the speed is within the thresholds set. This information leads to assignment of an SI_STATE value.

```{r check results of gear analysis, include=FALSE}
  # Assign for visually inspected gears a simple speed rule classification
  
  metiers <- unique(nonsubTacsat$LE_GEAR) # retrieve the unique gear types from the nonsubTacsat dataset
  nonsubTacsat$SI_STATE <- NA # set all entries in the SI_STATE column of the dataset to NA
  for (mm in metiers) { # for every unique gear type, complete the following
    nonsubTacsat$SI_STATE[
      nonsubTacsat$LE_GEAR == mm &
        nonsubTacsat$SI_SP >= speedarr[speedarr$LE_GEAR == mm, "min"] &
        nonsubTacsat$SI_SP <= speedarr[speedarr$LE_GEAR == mm, "max"]
    ] <- "f"; # if the gear type is in the metiers dataset, and the speed falls within the thresholds set, set the SI_STATE value to f
  }
  nonsubTacsat$SI_STATE[
    nonsubTacsat$LE_GEAR == "NA" &
      nonsubTacsat$SI_SP >= speedarr[speedarr$LE_GEAR == "MIS", "min"] &
      nonsubTacsat$SI_SP <= speedarr[speedarr$LE_GEAR == "MIS", "max"]
  ] <- "f"
  # if the gear type is not in the metiers dataset, and the speed does not fall within the thresholds set, set the SI_STATE value to s
  nonsubTacsat$SI_STATE[ is.na(nonsubTacsat$SI_STATE) ] <- "s"
  
  
  # Combine the two dataset together again 
  
  tacsatp <- rbindTacsat(subTacsat, nonsubTacsat) # rbind merges the datasets vertically
  tacsatp <- orderBy( ~ VE_REF + SI_DATIM, data = tacsatp) # sort the rows based on these columns
  
  
  # Set fishing sequences with hauling in the middle to "f" 
  
  idx <-
    which(
      tacsatp$SI_STATE[2:(nrow(tacsatp) - 1)] == "h" &
        tacsatp$SI_STATE[1:(nrow(tacsatp) - 2)] == "f" &
        tacsatp$SI_STATE[3:(nrow(tacsatp))    ] == "f" &
        tacsatp$VE_REF[2:(nrow(tacsatp) - 1)] == tacsatp$VE_REF[1:(nrow(tacsatp) - 2)] &
        tacsatp$VE_REF[2:(nrow(tacsatp) - 1)] == tacsatp$VE_REF[3:(nrow(tacsatp))]
    ) + 1
  tacsatp$SI_STATE[idx] <- "f"
  
  
  # save the outputs
  save(
    tacsatp,
    file = file.path(outPath, paste0("tacsatActivity", year, ".RData"))
  )
  
  message("Defining activity completed")
  
  
  # 2.3 Dispatch landings of merged eflalo at the ping scale
  
  idxkgeur <- kgeur(colnames(eflalo))
  eflalo$LE_KG_TOT <- rowSums(eflalo[,grep("LE_KG_",colnames(eflalo))],na.rm=T)
  eflalo$LE_EURO_TOT <- rowSums(eflalo[,grep("LE_EURO_",colnames(eflalo))],na.rm=T)
  eflalo <- eflalo[, -idxkgeur]
  eflaloNM <- subset(eflalo,!FT_REF %in% unique(tacsatp$FT_REF))
  eflaloM <- subset(eflalo,FT_REF %in% unique(tacsatp$FT_REF))
  
  tacsatp$SI_STATE[which(tacsatp$SI_STATE != "f")] <- 0
  tacsatp$SI_STATE[which(tacsatp$SI_STATE == "f")] <- 1
  
  tacsatEflalo <- tacsatp[tacsatp$SI_STATE == 1,] 
  
  
  # There are several options, specify at the top of this script what type of linking you require
  
  if (!"trip" %in% linkEflaloTacsat) stop("trip must be in linkEflaloTacsat")
  if (all(c("day", "ICESrectangle", "trip") %in% linkEflaloTacsat)) {
    tacsatEflalo <-
      splitAmongPings(
        tacsat = tacsatp,
        eflalo = eflaloM,
        variable = "all",
        level = "day",
        conserve = TRUE
      )
  } else
  {
    if (
      all(c("day","trip") %in% linkEflaloTacsat) &
      !"ICESrectangle" %in% linkEflaloTacsat
    ) {
      tmpTa <- tacsatp
      tmpEf <- eflaloM
      tmpTa$LE_RECT <- "ALL"
      tmpEf$LE_RECT <- "ALL"
      tacsatEflalo <-
        splitAmongPings(
          tacsat = tmpTa,
          eflalo = tmpEf,
          variable = "all",
          level = "day",
          conserve = TRUE
        )
    } else
    {
      if (
        all(c("ICESrectangle", "trip") %in% linkEflaloTacsat) &
        !"day" %in% linkEflaloTacsat
      )
      {
        tacsatEflalo <-
          splitAmongPings(
            tacsat = tacsatp,
            eflalo = eflaloM,
            variable = "all",
            level = "ICESrectangle",
            conserve = TRUE
          )
      } else
      {
        if (linkEflaloTacsat == "trip" & length(linkEflaloTacsat) == 1)
        {
          tacsatEflalo <-
            splitAmongPings(
              tacsat = tacsatp,
              eflalo = eflaloM,
              variable = "all",
              level = "trip",
              conserve = FALSE
            )
        }
      }
    }
  }
  
  save(
    tacsatEflalo,
    file = file.path(outPath, paste0("tacsatEflalo", year, ".RData"))
  )
  
  print("Dispatching landings completed")
  
  
  # 2.4 Assign c-square, year, month, quarter, area and create table 1 
  
  
  tacsatEflalo$Csquare   <- CSquare(tacsatEflalo$SI_LONG, tacsatEflalo$SI_LATI, degrees = 0.05)
  tacsatEflalo$Year      <- year(tacsatEflalo$SI_DATIM)
  tacsatEflalo$Month     <- month(tacsatEflalo$SI_DATIM)
  tacsatEflalo$kwHour    <- tacsatEflalo$VE_KW * tacsatEflalo$INTV / 60
  tacsatEflalo$INTV      <- tacsatEflalo$INTV / 60
  
  
  RecordType <- "VE"
  
  if(year == yearsToSubmit[1]) {
    table1 <-
      cbind(
        RT = RecordType,
        tacsatEflalo[,
                     c(
                       "VE_REF", "VE_COU", "Year", "Month", "Csquare", "LE_GEAR",
                       "LE_MET", "SI_SP", "INTV", "VE_LEN", "kwHour", "VE_KW", "LE_KG_TOT", "LE_EURO_TOT"
                     )
        ])
  } else {
    
    table1 <-
      rbind(
        table1,
        cbind(
          RT = RecordType,
          tacsatEflalo[,
                       c(
                         "VE_REF", "VE_COU", "Year", "Month", "Csquare", "LE_GEAR",
                         "LE_MET", "SI_SP", "INTV", "VE_LEN", "kwHour", "VE_KW", "LE_KG_TOT", "LE_EURO_TOT"
                       )
          ])
      )
    
  }
  
  
  # Save table1
  
  
  
  save(
    table1,
    file = file.path(outPath, "table1.RData" )
  )
  
  message(glue ("Table 1 for year {year} is completed") )
  
  
  # 2.5 Assign  year, month, quarter, area and create table 2
  
  eflalo$Year <- year(eflalo$FT_LDATIM) # retrieve the year from the FT_LDATIM column
  eflalo$Month <- month(eflalo$FT_LDATIM) # retrieve the month
  eflalo$INTV <- 1 # 1 day
  eflalo$dummy <- 1
  res <-
    aggregate(
      eflalo$dummy,
      by = as.list(eflalo[, c("VE_COU", "VE_REF", "LE_CDAT")]),
      FUN = sum,
      na.rm <- TRUE
    )
  colnames(res) <- c("VE_COU", "VE_REF", "LE_CDAT", "nrRecords")
  eflalo <- merge(eflalo, res, by = c("VE_COU", "VE_REF", "LE_CDAT"))
  eflalo$INTV <- eflalo$INTV / eflalo$nrRecords
  eflalo$kwDays <- eflalo$VE_KW * eflalo$INTV # multiply kw and intv to calculate kwdays of effort 
  eflalo$tripInTacsat <- ifelse(eflalo$FT_REF %in% tacsatp$FT_REF, "Y", "N") # Y = Yes and N = No
  
  
  
  RecordType <- "LE"
  
  if (year == yearsToSubmit[1]) {
  
    # if the year matches the entry, horizontally bind the quoted column names to the table2 variable
    table2 <-
      cbind(
        RT = RecordType,
        eflalo[
          ,
          c( # a list of the column headers being attached 
            "VE_REF", "VE_COU", "Year", "Month", "LE_RECT", "LE_GEAR", "LE_MET",
            "VE_LEN", "tripInTacsat", "INTV", "kwDays", "LE_KG_TOT", "LE_EURO_TOT"
          )
        ]
      )
    
  } else {
    
    # if the prior criteria was not met, the following will run
    table2 <-
      rbind(
        table2,
        cbind(
          RT = RecordType,
          eflalo[
            ,
            c( # a list of the column headers being attached 
              "VE_REF", "VE_COU", "Year", "Month", "LE_RECT", "LE_GEAR", "LE_MET",
              "VE_LEN", "tripInTacsat", "INTV", "kwDays", "LE_KG_TOT", "LE_EURO_TOT"
            )
          ]
        )
      )
    
  }
  
  
  
  # Save table2
  
  save(
    table2,
    file = file.path(outPath, "table2.RData" )
  )
  
  message(glue ("Table 2 for year {year} is completed") )

```


