---
title: "3_fisheries_data_products_SRF_2"
author: "Roi & Mike"
date: "2023-12-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(vmstools)
library(sf)
library(rnaturalearth)
library(devtools)
library(data.table)
library(here)

setwd(here(".\\data"))
```

Next the directories which will be used to save outputs in must be created. The dir.create() function will not cause an error if the folder already exists, so there is no need to use an ifelse statement. Each folder is created using the outPath, and then a subfolder added using the pate0() function.

```{r create paths and folders, include=TRUE, eval=FALSE}
# create folders for outputs if they do not exist 

# input folder
inPath = here(".\\analysis_toolbox\\3_fisheries_data_products\\fisheries_data_products_requests")

# data products folder
dir.create(paste0(inPath, "\\data-products"))
outPath = paste0(inPath, "\\data-products\\")

# geojson output folder
dir.create(paste0(outPath, "geojson"))
gjPath = paste0(outPath, "geojson\\")

# plots folder
dir.create(paste0(outPath, "plots"))
plotsPath = paste0(outPath, "plots\\")
```

Set the C-square size. This is just for file saving purposes. To change the outputs, the variable can be changed everywhere in the script using ctrl+f search and replace.

```{r set csquare size, include=TRUE, eval=FALSE}
csq = 'csq005'
```

Data product A calls for Welsh vessels of all lengths, wherever they are fishing. Therefore the Welsh fleet data is loaded. As such, the year is set to 2022. This could be changed to suit the requirements of a different data product in the future. The outputs of the workflow up to this point, for 2022, are loaded.

```{r output a setup, include=TRUE, eval=FALSE}
year = 2022

load(file = paste0(".\\workflow_outputs\\srf-2\\tacsatEflalo_welsh_fleet_", year, "_SRF_2_FDP.RData"))
load(file = paste0(".\\workflow_outputs\\srf-2\\eflalo_output_welsh_fleet_", year, "_SRF_2_FDP.RData"))
```

Adjusting the spatial resolution can also be an effective way of viewing the data. Here, a new column is added which will assign each row a csquare of 0.01 degrees. Previously, and most commonly, 0.05 degree csquares have been used in this analysis.

```{r 0.01 csquares, include=TRUE, eval=FALSE}
tacsatEflalo$Csquare_01 = CSquare(tacsatEflalo$SI_LONG, tacsatEflalo$SI_LATI, degrees = 0.01)
```

A unique vessel id column is necessary, so it is created here. Join the new data frame to the tacsatEflalo data in preparation for the data products.

```{r vessel id, include=TRUE, eval=FALSE}
VE_lut <- data.frame(VE_REF = unique(c(tacsatEflalo$VE_REF, eflalo_output$VE_REF)))
fmt <- paste0("%0", floor(log10(nrow(VE_lut))) + 1, "d")
VE_lut$VE_ID <- paste0(tacsatEflalo$VE_COU[1], sprintf(fmt, 1:nrow(VE_lut))) # use relevant country code!


# join onto data tables
table1 <- left_join(tacsatEflalo, VE_lut)
table1$VE_KW = as.numeric(table1$VE_KW)
```

Now the data must be pivoted into long format, with the individual species columns being converted into the LE_SPE column.

```{r long data format, include=TRUE, eval=FALSE}
table1 <- data.table(table1)

table1m <- melt(table1, 
            measure = patterns("LE_KG" = "LE_KG", "LE_EURO" = "LE_EURO"),
            value.name = c("LE_KG", "LE_EURO"), 
            variable.name = "LE_SPE")
```

Adjust the newly created LE_SPE column to include the species of each entry.

```{r convert species codes, include=TRUE, eval=FALSE}
table1m[, LE_SPE := factor(LE_SPE, 
                        labels = unique(gsub("(.*)(_)([A-Z]*)$", "\\3", 
                                             grep("LE_KG|LE_EURO", colnames(table1), value = TRUE))))]
```

Filter to entries where the weight or value is not 0.

```{r filter , include=TRUE, eval=FALSE}
table1m = table1m %>% filter(LE_KG > 0 & LE_EURO > 0)

table1m %>% filter(!INTV == 0) %>% tally()
```


###Data Product Creation

The data products created take the data and aggregate it based on certain categories within the dataset, such as spatial aggregation by csquare, or temporal aggregation by year or month. This helps to provide an overall view of the data over the time period or across the aoi. Additionally, categories, such as those created earlier, like vessel length, can show patterns.

Data product 1 is grouped by month, 0.05 degree Csquare, gear type and species.

```{r data product a, include=TRUE, eval=FALSE}
### Year, Month, 0.05 Csquare, gear, species

table1.a_input <- table1m

table1.a =
  table1.a_input %>%
  group_by(SI_YEAR, Month, Csquare_05, LE_GEAR, LE_SPE) %>%
  summarise(
    mean_si_sp       = mean(SI_SP),
    sum_intv         = sum(INTV, na.rm = TRUE),
    mean_intv        = mean(INTV, na.rm=TRUE),
    mean_ve_len      = mean(VE_LEN, na.rm = TRUE),
    mean_ve_kw       = mean(VE_KW, na.rm = TRUE),
    sum_kwHour       = sum(kwHour, na.rm=TRUE),
    #sum_le_kg_tot    = sum(LE_KG_TOT, na.rm = TRUE),
    #sum_le_euro_tot  = sum(LE_EURO_TOT, na.rm = TRUE),
    sum_le_kg_tot    = sum(LE_KG, na.rm = TRUE),
    sum_le_euro_tot  = sum(LE_EURO, na.rm = TRUE),
    n_vessels        = n_distinct(VE_ID, na.rm = TRUE),
    vessel_ids       = ifelse (
      n_distinct(VE_ID) < 3,
      paste(unique(VE_ID), collapse = ";"),
      'not_required'
    )
  ) %>% relocate(n_vessels, vessel_ids, .after = LE_GEAR) %>%
  #mutate(AverageGearWidth = NA %>% as.numeric()) %>%
  as.data.frame()

write.csv(table1.a, paste0(outPath, "\\", year, "_table1_a_", csq, "_SRF2.csv"), row.names = FALSE)
```



The above processes are repeated, but for the other data products. Their code descriptions are not included, as the code is the same as above.

This is data product B, which analyses all vessels fishing in Welsh waters, for 2022.

```{r data produt b, include=TRUE, eval=FALSE}
year = 2022

load(file = paste0(".\\workflow_outputs\\srf-2\\tacsatEflalo_welsh_waters_", year, "_SRF_2_FDP.RData"))
load(file = paste0(".\\workflow_outputs\\srf-2\\eflalo_output_welsh_waters_", year, "_SRF_2_FDP.RData"))

## ASSGIN C-SQuare 0.01
tacsatEflalo$Csquare_01 = CSquare(tacsatEflalo$SI_LONG, tacsatEflalo$SI_LATI, degrees = 0.01)

VE_lut <- data.frame(VE_REF = unique(c(tacsatEflalo$VE_REF, eflalo_output$VE_REF)))
fmt <- paste0("%0", floor(log10(nrow(VE_lut))) + 1, "d")
VE_lut$VE_ID <- paste0(tacsatEflalo$VE_COU[1], sprintf(fmt, 1:nrow(VE_lut))) # use relevant country code!

# join onto data tables
table1 <- left_join(tacsatEflalo, VE_lut)
table1$VE_KW = as.numeric(table1$VE_KW)

table1 <- data.table(table1)

table1m <- melt(table1, 
                measure = patterns("LE_KG" = "LE_KG", "LE_EURO" = "LE_EURO"),
                value.name = c("LE_KG", "LE_EURO"), 
                variable.name = "LE_SPE")

table1m[, LE_SPE := factor(LE_SPE, 
                           labels = unique(gsub("(.*)(_)([A-Z]*)$", "\\3", 
                                                grep("LE_KG|LE_EURO", colnames(table1), value = TRUE))))]

table1m = table1m %>% filter(LE_KG > 0 & LE_EURO > 0)

table1.b_input <- table1m

### Year, Month, 0.05 Csquare, gear, species
table1.b = 
  table1.b_input %>%
  group_by(SI_YEAR, Month, Csquare_05, LE_GEAR, LE_SPE) %>%
  summarise(
    mean_si_sp       = mean(SI_SP),
    sum_intv         = sum(INTV, na.rm = TRUE),
    mean_intv        = mean(INTV, na.rm=TRUE),
    mean_ve_len      = mean(VE_LEN, na.rm = TRUE),
    mean_ve_kf       = mean(VE_KW, na.rm = TRUE),
    sum_kwHour       = sum(kwHour, na.rm=TRUE),
    sum_le_kg        = sum(LE_KG, na.rm = TRUE),
    sum_le_euro      = sum(LE_EURO, na.rm = TRUE),
    n_vessels        = n_distinct(VE_ID, na.rm = TRUE),
    vessel_ids       = ifelse (
      n_distinct(VE_ID) < 3,
      paste(unique(VE_ID), collapse = ";"),
      'not_required'
    )
  ) %>% relocate(n_vessels, vessel_ids, .before = Csquare_05) %>%
  mutate(AverageGearWidth = NA %>% as.numeric()) %>%
  as.data.frame()

write.csv(table1.b, paste0(outPath, "\\", year, "_table1_b_", csq, "_SRF2.csv"), row.names = FALSE)
```


Finally, for data products C and D, multiple years are called for, so a for loop is used. The years range can be changed in the first section, it can be set to an infinite range of years, limited by the data, or just one year. For this product, the analysis type must be chosen, as it dictates which data is loaded and the name of the file output.

```{r analysis set up, include=TRUE, eval=FALSE}
years = 2012:2021
# year = 2021
analysis_type = 'welsh_waters' ## welsh_fleet / welsh_waters
```

To create a large table of all the years combined, bind_rows must be used. As 2012 is the first year, it has nothing to be bound to, so must just be written to the binding object by itself. By using an if statement, this can be managed, with every year after 2012 being bound to the object subsequently.

```{r data products c/d, include=TRUE, eval=FALSE}
for ( year in years ) {
  
  load(file = paste0(".\\workflow_outputs\\srf-2\\tacsatEflalo_", analysis_type, "_", year, "_SRF_2_FDP.RData"))
  load(file = paste0(".\\workflow_outputs\\srf-2\\eflalo_output_", analysis_type, "_", year, "_SRF_2_FDP.RData"))
  
  if (year == 2012) {
    tacsatEflalo_all <- tacsatEflalo
    eflalo_all <- eflalo_output
  } else {
    tacsatEflalo_all <- tacsatEflalo_all %>% bind_rows(., tacsatEflalo)
    eflalo_all <- eflalo_all %>% bind_rows(., eflalo_output)
  }
} 

  tacsatEflalo <- tacsatEflalo_all
  eflalo_output <- eflalo_all
  
  tacsatEflalo$Csquare_01 = CSquare(tacsatEflalo$SI_LONG, tacsatEflalo$SI_LATI, degrees = 0.01)
  
  VE_lut <- data.frame(VE_REF = unique(c(tacsatEflalo$VE_REF, eflalo_output$VE_REF)))
  fmt <- paste0("%0", floor(log10(nrow(VE_lut))) + 1, "d")
  VE_lut$VE_ID <- paste0(tacsatEflalo$VE_COU[1], sprintf(fmt, 1:nrow(VE_lut))) # use relevant country code!
  
  
  # join onto data tables
  table1 <- left_join(tacsatEflalo, VE_lut)
  table1$VE_KW = as.numeric(table1$VE_KW)
  
  table1 <- data.table(table1)
  
  table1m <- melt(table1, 
                  measure = patterns("LE_KG" = "LE_KG", "LE_EURO" = "LE_EURO"),
                  value.name = c("LE_KG", "LE_EURO"), 
                  variable.name = "LE_SPE")
  
  table1m[, LE_SPE := factor(LE_SPE, 
                             labels = unique(gsub("(.*)(_)([A-Z]*)$", "\\3", 
                                                  grep("LE_KG|LE_EURO", colnames(table1), value = TRUE))))]
  
  table1m = table1m %>% filter(LE_KG > 0 & LE_EURO > 0)
  
  table1_input = table1m %>% filter (VE_LEN > 12)
  
  ### Year, Month, 0.05 Csquare, gear, species
  table1.output = 
    table1_input %>%
    group_by(SI_YEAR, Month, Csquare_05, LE_GEAR, LE_SPE) %>%
    summarise(
      mean_si_sp       = mean(SI_SP),
      sum_intv         = sum(INTV, na.rm = TRUE),
      mean_intv        = mean(INTV, na.rm=TRUE),
      mean_ve_len      = mean(VE_LEN, na.rm = TRUE),
      mean_ve_kw       = mean(VE_KW, na.rm = TRUE),
      sum_kwHour       = sum(kwHour, na.rm=TRUE),
      sum_le_kg        = sum(LE_KG, na.rm = TRUE),
      sum_le_euro      = sum(LE_EURO, na.rm = TRUE),
      n_vessels        = n_distinct(VE_ID, na.rm = TRUE),
      vessel_ids       = ifelse (
        n_distinct(VE_ID) < 3,
        paste(unique(VE_ID), collapse = ";"),
        'not_required'
      )
    ) %>% relocate(n_vessels, vessel_ids, .after = LE_GEAR) %>%
    #mutate(AverageGearWidth = NA %>% as.numeric()) %>%
    as.data.frame()
  
  # Use an if statement to dictate the output file names, based on the analysis type specified prior to the loop
  
  if (analysis_type == 'welsh_fleet') {
    write.csv(table1.output, paste0(outPath, "\\table1_c_", csq, "_SRF2.csv"), row.names = FALSE)
  } else if (analysis_type == 'welsh_waters') {
    write.csv(table1.output, paste0(outPath, "\\table1_d_", csq, "_SRF2.csv"), row.names = FALSE)
  }
```