---
title: "3_fisheries_data_products"
author: "Roi & Mike"
date: "03/04/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lubridate)
library(dplyr)
library(sf)
library(ggplot2)
library(vmstools)
```

```{r load data, include=FALSE}
# 3.1 Load TABLE 1 (VMS) and TABLE 2 (LOGBOOK) --------------------------------------------
#  Use the load function to bring in the tables made in the previous workflow
load(file = paste0(outPath, "table1.RData")  )
load( file = paste0(outPath, "table2.RData")  )
```


```{r combine tables 1 and 2, include=FALSE}
# 3.2 Replace vessel id by an anonymised id column  --------------------------------------------

# New field added for the 2020 datacall including unique vessels id's  #
# This vessel id is used to calculate unique vessels in a c-square and  #

# create a dataframe containing the unique vessel references from tables 1 and 2
VE_lut <- data.frame(VE_REF = unique(c(table1$VE_REF, table2$VE_REF)))

# retrieve the log10 value of the number of rows in the VE_lut dataframe. Paste this between the contents
# of the two "" sections.
fmt <- paste0("%0", floor(log10(nrow(VE_lut))) + 1, "d")

# create the column VE_ID in the VE_lut dataset
VE_lut$VE_ID <- paste0(table1$VE_COU[1], sprintf(fmt, 1:nrow(VE_lut))) # use relevant country code!


# join onto data tables, maintaining all the data from tables 1 and 2 and applying matches from the VE_lut dataset
table1 <- left_join(table1, VE_lut)
table2 <- left_join(table2, VE_lut)
```



```{r create length categories, include=FALSE}
# 3.3 Assign the vessel length category based in DATSU vocabulary   ----------------------


#  Use of the "icesVocab" ICES developed R package that fetch the DATSU vocabulary values for a given vocabulary theme #
# install.packages("icesVocab", repos = "https://ices-tools-prod.r-universe.dev")

library(icesVocab)

# Get the values accepted in this vocabulary dataset

vlen_ices <- getCodeList("VesselLengthClass") ### Get DATSU Vocabulary list for selected dataset


# Filter values that aren't deprecated, overlapped or not accepted by data call requirements

vlen_icesc <-  vlen_ices%>%slice(2, 4, 6, 7, 8, 10, 11, 12, 13 )%>%select(Key)


# TABLE 1. Add the vessel length category using  LENGTHCAT field

# this section of code uses the breaks function to categorise the length of each vessel. Categories are designated by the numbers within the brackets.
table1$LENGTHCAT <-  table1$VE_LEN%>%cut(    breaks=c(0, 6, 8, 10, 12, 15, 18, 24, 40, 'inf' ), 
                                             right = FALSE    ,include.lowest = TRUE,
                                             labels =  vlen_icesc$Key 
)


# TABLE 2. Add the vessel length category using  LENGTHCAT field, same as for table 1.

table2$LENGTHCAT <-  table2$VE_LEN%>%cut(   breaks=c(0, 6, 8, 10, 12, 15, 18, 24, 40, 'inf' ), 
                                            right = FALSE    ,include.lowest = TRUE,
                                            labels =  vlen_icesc$Key 
)
```

Change the formatting of tables 1 and 2 to show summary statistics, rather than raw data.

```{r aggregate and summarise, include=FALSE}
# 3.4 Aggregate and summarise TABLE 1 and TABLE2   ----------------------

# Table 1

table1Save <-
  table1 %>% # the main input into the new variable is table 1
  separate(col = LE_MET ,   c("met4", "met5" ), sep = '_', remove = FALSE)%>% # separate the LE_MET column into metier levels 4 and 5, this is done via the underscore
  group_by(RT,VE_COU,Year,Month,Csquare,LE_GEAR, met5,  LE_MET,LENGTHCAT) %>% # group the data together by these chosen categories
  summarise(
    mean_si_sp       = mean(SI_SP), # calculate the mean of the SI_SP column for each trip 
    sum_intv         = sum(INTV, na.rm=TRUE), # sum the INTV
    mean_intv        = mean(INTV, na.rm=TRUE), # calculate the mean INTV
    mean_ve_len      = mean(VE_LEN, na.rm = TRUE), # calculate the mean vessel length 
    mean_ve_kf       = mean(VE_KW, na.rm = TRUE), 
    sum_kwHour       = sum(kwHour, na.rm=TRUE), # total Kwh
    sum_le_kg_tot    = sum(LE_KG_TOT, na.rm = TRUE), # total weight
    sum_le_euro_tot  = sum(LE_EURO_TOT, na.rm = TRUE), # total value
    n_vessels        = n_distinct(VE_ID, na.rm = TRUE),
    vessel_ids       = ifelse (
      n_distinct(VE_ID) < 3,
      paste(unique(VE_ID), collapse = ";"),
      'not_required'
    )
  ) %>%  relocate( n_vessels,vessel_ids, .before = Csquare)%>%
  mutate (AverageGearWidth = NA%>%as.numeric()  )%>% ## If this information is available modify this line of the script. By default is assumed not existing gear width information
  as.data.frame()

# set the column headings for the table1Save table to the following
colnames(table1Save) <-
  c(
    "RecordType", "CountryCode", "Year", "Month", "NoDistinctVessels", "AnonymizedVesselID",
    "C-square","MetierL4", "MetierL5",  "MetierL6",  "VesselLengthRange",
    "AverageFishingSpeed", "FishingHour", "AverageInterval", "AverageVesselLength", "AveragekW",
    "kWFishingHour", "TotWeight", "TotValue" , "AverageGearWidth"
  )




# Table 2

table2Save <-
  table2 %>%
  separate( col = LE_MET ,   c("met4", "met5"  ), sep = '_', remove = FALSE)%>% # separate the LE_MET column into metier levels 4 and 5, this is done via the underscore
  group_by( RT, VE_COU, Year, Month, LE_RECT,LE_GEAR, met5,  LE_MET,LENGTHCAT, tripInTacsat ) %>%
  summarise(
    sum_intv        = sum(INTV, na.rm = TRUE), # sum the INTV
    sum_kwDays      = sum(kwDays, na.rm = TRUE), # calculate the effort in kilowatt days
    sum_le_kg_tot   = sum(LE_KG_TOT, na.rm = TRUE), # total weight
    sum_le_euro_tot = sum(LE_EURO_TOT, na.rm = TRUE), # total value
    n_vessels       = n_distinct(VE_ID, na.rm = TRUE),
    vessel_ids      = ifelse (
      n_distinct(VE_ID) < 3,
      paste(
        unique(VE_ID), collapse = ";"),
      'not_required'
    )
  ) %>%  relocate( n_vessels,vessel_ids, .before = LE_RECT)%>%
  as.data.frame()

# set the column headings for the table2Save table to the following
colnames(table2Save) <-
  c(
    "RecordType", "CountryCode", "Year", "Month", "NoDistinctVessels", "AnonymizedVesselID", "ICESrectangle",
    "MetierL4", "MetierL5",   "MetierL6", "VesselLengthRange", "VMSEnabled", "FishingDays",
    "kWFishingDays", "TotWeight", "TotValue"
  )
```

The following sections run the tables through some quality checks to ensure the data included is limited to what is required. It will also discover any errors, such as the number of fishing hours being 0 by mistake, so that errors can be addressed.

```{r quality checks 1, include=FALSE}
# 3.5   ICES DATSU VOCABULARY CHECKS BEFORE DATA SUBMISSION  ------------------------------------------


##Get vocabulary for mandatory and fields with associated vocabulary using the DATSU API
# install.packages("icesVocab", repos = "https://ices-tools-prod.r-universe.dev")
library(icesVocab)


# TABLE 1 =============================================================


### 3.5.1 Check if C-Squares are within ICES Ecoregions =====================

csquares_d      <-  table1Save%>% # retrieve the unique csquares from table 1
  select('C-square')%>%
  distinct( ) # removes duplicates

# create a dataframe which houses the 0.05 degree csquares and the long/lat coordinates 
csquares_dcoord <-  cbind ( csquares_d ,  CSquare2LonLat (csqr = csquares_d$`C-square` ,degrees =  0.05)   )

# retrieve the csquares with latitude coordinates between 30 and 90 and apply them to the variable valid_csquare
valid_csquare   <-  csquares_dcoord%>%
  filter(SI_LATI >= 30 & SI_LATI <= 90  )%>%
  select('C-square')%>%
  pull()


# filter table 1 to only include entries which have common csquares with the valid_csquare table
table1Save      <-  table1Save%>%filter(`C-square` %in% valid_csquare)



### 3.5.2 Check Vessel Lengths categories are accepted ==================================


vlen_ices       <-  getCodeList("VesselLengthClass")
table ( table1Save$VesselLengthRange%in%vlen_ices$Key )  # TRUE records accepted in DATSU, FALSE aren't

# Get summary  of   DATSU valid/not valid records
table1Save [ !table1Save$VesselLengthRange %in%vlen_ices$Key,]%>%group_by(VesselLengthRange)%>%select(VesselLengthRange)%>%tally()

# Correct them if any not valid and filter only valid ones
table1Save      <-  table1Save%>%filter(VesselLengthRange %in% vlen_ices$Key)



### 3.5.3 Check Metier L4 (Gear) categories are accepted =================================

m4_ices         <-  getCodeList("GearTypeL4")
table (table1Save$MetierL4 %in%m4_ices$Key )   # TRUE records accepted in DATSU, FALSE aren't

# Get summary  of   DATSU valid/not valid records
table1Save [ !table1Save$MetierL4 %in%m4_ices$Key,]%>%group_by(MetierL4)%>%select(MetierL4)%>%tally()

# Correct them if any not valid and filter only valid ones
table1Save      <-  table1Save%>%filter(MetierL4 %in% m4_ices$Key)


### 3.5.4 Check Metier L5 (Target Assemblage) categories are accepted =====================

m5_ices         <-  getCodeList("TargetAssemblage")

table (table1Save$MetierL5 %in%m5_ices$Key )   # TRUE records accepted in DATSU, FALSE aren't

# Get summary  of   DATSU valid/not valid records
table1Save [ !table1Save$MetierL5 %in%m5_ices$Key,]%>%group_by(MetierL5)%>%select(MetierL5)%>%tally()

# Correct them if any not valid and filter only valid ones
table1Save      <-  table1Save%>%filter(MetierL5 %in% m5_ices$Key)

### 3.5.5 Check Metier L6 (Fishing Activity) categories are accepted =====================

m6_ices         <-  getCodeList("Metier6_FishingActivity")

table (table1Save$MetierL6 %in%m6_ices$Key )   # TRUE records accepted in DATSU, FALSE aren't

# Get summary  of   DATSU valid/not valid records
table1Save [ !table1Save$MetierL6 %in%m6_ices$Key,]%>%group_by(MetierL6)%>%select(MetierL6)%>%tally()

# Correct them if any not valid and filter only valid ones
table1Save      <-  table1Save%>%filter(MetierL6 %in% m6_ices$Key)

### 3.5.6 Check country codes =====================

cntrcode <- getCodeList("ISO_3166")
table (table1Save$CountryCode %in%cntrcode$Key )   # TRUE records accepted in DATSU, FALSE aren't

# Get summary  of   DATSU valid/not valid records
table1Save [ !table1Save$VMSEnabled %in% cntrcode$Key,]%>% group_by(CountryCode) %>% select(CountryCode) %>% tally()

# Correct them if any not valid and filter only valid ones
table1Save      <-  table1Save%>%filter(CountryCode %in% cntrcode$Key)

```

Repeat the above checks for table 2.

```{r quality checks 2, include=FALSE}
# TABLE 2  =============================================================


### 3.5.6 Check ICES rect are valid  =====================

statrect_ices <- getCodeList("StatRec")

table (table2Save$ICESrectangle %in%statrect_ices$Key )   # TRUE records accepted in DATSU, FALSE aren't

# Get summary  of   DATSU valid/not valid records
table2Save [ !table2Save$ICESrectangle %in%statrect_ices$Key,]%>%group_by(ICESrectangle)%>%select(ICESrectangle)%>%tally()

# Correct them if any not valid and filter only valid ones
table2Save      <-  table2Save%>%filter(ICESrectangle %in% statrect_ices$Key)



### 3.5.7 Check Vessel Lengths categories are accepted ==================================


vlen_ices       <-  getCodeList("VesselLengthClass")
table ( table2Save$VesselLengthRange%in%vlen_ices$Key )  # TRUE records accepted in DATSU, FALSE aren't

# Get summary  of   DATSU valid/not valid records
table2Save [ !table2Save$VesselLengthRange %in%vlen_ices$Key,]%>%group_by(VesselLengthRange)%>%select(VesselLengthRange)%>%tally()

# Correct them if any not valid and filter only valid ones
table2Save      <-  table2Save%>%filter(VesselLengthRange %in% vlen_ices$Key)


### 3.5.8 Check Metier L4 (Gear) categories are accepted =================================

m4_ices         <-  getCodeList("GearTypeL4")
table (table2Save$MetierL4 %in%m4_ices$Key )   # TRUE records accepted in DATSU, FALSE aren't

# Get summary  of   DATSU valid/not valid records
table2Save [ !table2Save$MetierL4 %in%m4_ices$Key,]%>%group_by(MetierL4)%>%select(MetierL4)%>%tally()

# Correct them if any not valid and filter only valid ones
table2Save      <-  table2Save%>%filter(MetierL4 %in% m4_ices$Key)



### 3.5.9 Check Metier L5 (Target Assemblage) categories are accepted =====================

m5_ices         <-  getCodeList("TargetAssemblage")

table (table2Save$MetierL5 %in%m5_ices$Key )   # TRUE records accepted in DATSU, FALSE aren't

# Get summary  of   DATSU valid/not valid records
table2Save [ !table2Save$MetierL5 %in%m5_ices$Key,]%>%group_by(MetierL5)%>%select(MetierL5)%>%tally()

# Correct them if any not valid and filter only valid ones
table2Save      <-  table2Save%>%filter(MetierL5 %in% m5_ices$Key)



### 3.5.10 Check Metier L6 (Fishing Activity) categories are accepted =====================

m6_ices         <-  getCodeList("Metier6_FishingActivity")

table (table2Save$MetierL6 %in%m6_ices$Key )   # TRUE records accepted in DATSU, FALSE aren't

# Get summary  of   DATSU valid/not valid records
table2Save [ !table2Save$MetierL6 %in%m6_ices$Key,]%>%group_by(MetierL6)%>%select(MetierL6)%>%tally()

# Correct them if any not valid and filter only valid ones
table2Save      <-  table2Save%>%filter(MetierL6 %in% m6_ices$Key)



### 3.5.11 Check VMSEnabled categories are accepted =====================


yn <- getCodeList("YesNoFields")

table (table2Save$VMSEnabled %in%yn$Key )   # TRUE records accepted in DATSU, FALSE aren't

# Get summary  of   DATSU valid/not valid records
table2Save [ !table2Save$VMSEnabled %in%yn$Key,]%>%group_by(VMSEnabled)%>%select(VMSEnabled)%>%tally()

# Correct them if any not valid and filter only valid ones
table2Save      <-  table2Save%>%filter(VMSEnabled %in% yn$Key)



### 3.5.12 Check country codes =====================

cntrcode <- getCodeList("ISO_3166")
table (table2Save$CountryCode %in%cntrcode$Key )   # TRUE records accepted in DATSU, FALSE aren't

# Get summary  of   DATSU valid/not valid records
table2Save [ !table2Save$VMSEnabled %in% cntrcode$Key,]%>% group_by(CountryCode) %>% select(CountryCode) %>% tally()

# Correct them if any not valid and filter only valid ones
table2Save      <-  table2Save%>%filter(CountryCode %in% cntrcode$Key)


# DATSU Vocabulary check finished
```

Further, optional, quality checks.

```{r quality checks 3, include=FALSE}
# 3.6 DATA QC REPORT (OPTIONAL)   ------------------------------------------



# Null values are only accepted for NON MANDATORY fields

# TABLE 1 =============================================================

# Create the table to check fields formats and number of NA's

table_nas <- NULL
for ( nn in colnames(table1Save)) {
  table_na <- table(table1Save[, nn]%>%is.na() )
  row <- c(field = nn, is_na =  ifelse(is.na (table_na['TRUE']), 0, table_na['TRUE'] ), total_records =  table1Save[, nn]%>%length(), field_type =class(  table1Save[, nn]  ) )
  table_nas <- rbind(table_nas,  row)
}

# Print a summary table in Viewer

gt(
  table_nas%>%as_tibble(),
  rowname_col = 'field'
) %>%
  tab_header(
    title = md('Summary of **Table 1**  number of NA and records types')
  ) %>%
  cols_label(  `is_na.NA`=  md('Number of  <br> NA\'s') ,
               total_records = md('Total <br> records'),
               field_type = md('Field <br> type')
  ) %>%
  tab_footnote(
    footnote = md('Non mandatory fields can include null values if not available'),
    locations = cells_stub( rows = c( 'TotValue', 'AverageGearWidth')  )
  )


# TABLE 2 =============================================================

# Create the table to check fields formats and number of NA's

table_nas <- NULL
for ( nn in colnames(table2Save)) { # initiate a for loop where nn is the iterator through the column names of table2save
  table_na <- table(table2Save[, nn]%>%is.na() )
  row <- c(field = nn, is_na =  ifelse(is.na (table_na['TRUE']), 0, table_na['TRUE'] ), total_records =  table2Save[, nn]%>%length(), field_type =class(  table2Save[, nn]  ) )
  table_nas <- rbind(table_nas,  row)
}

# Print a summary table in Viewer

gt(
  table_nas%>%as_tibble(),
  rowname_col = 'field'
) %>%
  tab_header(
    title = md('Summary of **Table 2**  number of NA and records types')
  ) %>%
  cols_label(  `is_na.NA`=  md('Number of  <br> NA\'s') ,
               total_records = md('Total <br> records'),
               field_type = md('Field <br> type')
  ) %>%
  tab_footnote(
    footnote = md('Non mandatory fields can include null values if not available'),
    locations = cells_stub( rows = c( 'TotValue')  )
  )


# Check if TABLE 1 fishing hours > 0

table( table1$INTV > 0  )

# Check if TABLE 2 fishing days  > 0

table( table2$INTV > 0  )

# End of QC checks
```

Finally, save tables 1 and 2, these will be ready for datacall submission. Some final formatting, such as the removal of row numbers, are applied.

```{r save tables, include=FALSE}
# 3.7 Save the final TABLE 1 and TABLE 2 for datacall submission --------------------------------------------

## Headers and quotes have been removed to be compatible with required submission and ICES SQL DB format.

write.table(table1Save, file.path(outPath, "table1Save.csv"), na = "",row.names=FALSE,col.names=FALSE,sep=",",quote=FALSE)
write.table(table2Save, file.path(outPath, "table2Save.csv"), na = "",row.names=FALSE,col.names=FALSE,sep=",",quote=FALSE)




############### DATACALL SUBMISSION USING ICESVMS R PACKAGE (OPTIONAL)  ##################

# R packages required to be installed:
# install.packages(c("icesVMS", "icesConnect"), repos = "https://ices-tools-prod.r-universe.dev")  

library(icesVMS)

# Replace with your ICES user name and you will be requested with your password
icesConnect::set_username('submitter_ices_user_id') 

# icesConnect::ices_token(refresh = TRUE)
# icesConnect::decode_token()$UserEmail # Check the email associated to your ices user name is the correct one

screen_vms_file(file.path(outPath, "table1Save.csv"))  # Submit for screening Table 1
screen_vms_file(file.path(outPath, "table2Save.csv"))  # Submit for screening Table 2
```