---
title: "3_fisheries_data_products"
author: "Roi & Mike"
date: "03/04/2023"
output: html_document
---

```{r setup, include=TRUE, eval=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(dplyr)
library(sf)
library(vmstools)
```

It is common practice for functions to be placed at the top of a script, so here is a function which will be used later on. It will be evident when it is being used, as just like any other function, it will be called upon followed by opening brackets - plot_function(). This function is used to retrieve the elements of each table, to be used in plotting. Its required elements are in the first line

(data_plot, col_plot, col_facet1 = NULL, col_facet2 = NULL),

these are arbitrary names, which can be anything the user decides. They will be called upon in the function's creation, but the actual elements will only be chosen when the function it put to use later on.

```{r plotting function, include=TRUE, eval=FALSE}
# function for plotting data products - to be used later, it is just common practice to put functions at the top of scripts
plot_function = function (data_plot, col_plot, col_facet1 = NULL, col_facet2 = NULL) {
  
  gg1 =  ggplot() + geom_sf( data =   world     ) +
    geom_sf (data = data_plot , aes(color = sum_intv,  fill = sum_intv) ) +
    theme_bw() +
    coord_sf(xlim = lons, ylim = lats, expand = FALSE) #+
    #labs(fill = "Effort") # +
  # facet_wrap( ~ get( col_facet1) + get( col_facet2)  )
  if ( is.null( col_facet2 )  )  {
    gg1 =  gg1 +
      facet_wrap( ~ get( col_facet1)  )
  } else  {
    gg1 =  gg1 +
      facet_wrap( ~ get( col_facet1) +get( col_facet2)   )
  }
  return(gg1)
}
```

Next the directories which will be used to save outputs in must be created. The dir.create() function will not cause an error if the folder already exists, so there is no need to use an ifelse statement. Each folder is created using the outPath, and then a subfolder added using the pate0() function.

```{r filepaths, include=TRUE, eval=FALSE}
# input folder - this must be edited to where your workflow 2 outputs are located
inPath = "C:\\Users\\wg23\\rcourse\\2\\data_output"

# data products folder
dir.create("C:\\Users\\md09\\Documents\\git\\welsh_gov_fisheries_toolbox\\analysis_toolbox\\3_fisheries_data_products\\data-products")
outPath = "C:\\Users\\md09\\Documents\\git\\welsh_gov_fisheries_toolbox\\analysis_toolbox\\3_fisheries_data_products\\data-products\\"

# geojson output folder
dir.create(paste0(outPath, "geojson"))
gjPath = paste0(outPath, "geojson\\")

# plots folder
dir.create(paste0(outPath, "plots"))
plotsPath = paste0(outPath, "plots\\")
```

The year for which the analysis will be carried out.

```{r year, include=TRUE, eval=FALSE}
year = 2022
```

Load the files from the previous workflow.

```{r load files, include=TRUE, eval=FALSE}
load(file = paste0(inPath, "/eflalo_output_", year , ".RData"))
load(file = paste0(inPath, "/tacsatEflalo_output_", year , ".RData"))
```

Next, some additional columns will be added to the datasets which can be used for creating data products. The length category column helps to aggregate the data in different ways, reducing its granularity and providing a more general view of the data, using the specified categories.

```{r length category, include=TRUE, eval=FALSE}
# Create vessel length class: Add the vessel length category using  LENGTHCAT field
len_labels = c("0-5m", "5-7m", "7-10m", ">10m")

# assign a category to each entry based on the vessel length. Category limits are assigned as follows
tacsatEflalo$LENGTHCAT <- tacsatEflalo$VE_LEN %>% cut(breaks=c(0, 5, 7, 10, 'inf'), 
                                                      right = FALSE ,include.lowest = TRUE,
                                                      labels = len_labels
                                                      )
```

Similarly, by assigning a quarter based on the month column, data can be viewed per quarter. Differing temporal resolutions are useful to gain perspective of the data over different time spans.

```{r quarter, include=TRUE, eval=FALSE}
quar_labels = c("1", "2", "3", "4")
tacsatEflalo$QUARTER = tacsatEflalo$Month %>% cut(breaks = c(0, 4, 7, 10, 12),
                                                  right = F, include.lowest = T,
                                                  labels = quar_labels
                                                  )
```

Adjusting the spatial resolution can also be an effective way of viewing the data. Here, a new column is added which will assign each row a csquare of 0.01 degrees. Previously, and most commonly, 0.05 degree csquares have been used in this analysis.

```{r 0.01 csquares, include=TRUE, eval=FALSE}
tacsatEflalo$Csquare_01 = CSquare(tacsatEflalo$SI_LONG, tacsatEflalo$SI_LATI, degrees = 0.01)
```

A unique vessel id column is necessary, so it is created here.

```{r vessel id, include=TRUE, eval=FALSE}
VE_lut <- data.frame(VE_REF = unique(c(tacsatEflalo$VE_REF, eflalo_output$VE_REF)))
fmt <- paste0("%0", floor(log10(nrow(VE_lut))) + 1, "d")
VE_lut$VE_ID <- paste0(tacsatEflalo$VE_COU[1], sprintf(fmt, 1:nrow(VE_lut))) # use relevant country code!

table1 <- left_join(tacsatEflalo, VE_lut) # join onto data tables
```

###Data Product Creation

The data products created take the data and aggregate it based on certain categories within the dataset, such as spatial aggregation by csquare, or temporal aggregation by year or month. This helps to provide an overall view of the data over the time period or across the aoi. Additionally, categories, such as those created earlier, like vessel length, can show patterns.

Data product 1 is grouped by Quarter, 0.05 degree Csquare and gear type.

```{r data product 1, include=TRUE, eval=FALSE}
### Quarter, 0.05 Csquare, gear

table1.1 = 
  table1 %>%
  group_by(VE_COU,QUARTER,Csquare,LE_GEAR ) %>%
  summarise(
    mean_si_sp       = mean(SI_SP),
    sum_intv         = sum(INTV, na.rm = TRUE),
    mean_intv        = mean(INTV, na.rm=TRUE),
    mean_ve_len      = mean(VE_LEN, na.rm = TRUE),
    mean_ve_kw       = mean(VE_KW, na.rm = TRUE),
    sum_kwHour       = sum(kwHour, na.rm=TRUE),
    sum_le_kg_tot    = sum(LE_KG_TOT, na.rm = TRUE),
    #sum_le_euro_tot  = sum(LE_EURO_TOT, na.rm = TRUE),
    n_vessels        = n_distinct(VE_ID, na.rm = TRUE),
    vessel_ids       = ifelse (
      n_distinct(VE_ID) < 3,
      paste(unique(VE_ID), collapse = ";"),
      'not_required'
    )
  ) %>% relocate(n_vessels, vessel_ids, .before = Csquare) %>%
  #mutate(AverageGearWidth = NA %>% as.numeric()) %>%
  as.data.frame()

write.csv(table1.1, paste0(outPath, year, "_table1_1.csv"))
```

Data product 2 is grouped by Year, 0.05 degree Csquare and gear type.

```{r data product 2, include=TRUE, eval=FALSE}
### Year, 0.05 Csquare, gear
table1.2 = 
  table1 %>%
  group_by(VE_COU,Year,Csquare,LE_GEAR) %>%
  summarise(
    mean_si_sp       = mean(SI_SP),
    sum_intv         = sum(INTV, na.rm = TRUE),
    mean_intv        = mean(INTV, na.rm=TRUE),
    mean_ve_len      = mean(VE_LEN, na.rm = TRUE),
    mean_ve_kf       = mean(VE_KW, na.rm = TRUE),
    sum_kwHour       = sum(kwHour, na.rm=TRUE),
    sum_le_kg_tot    = sum(LE_KG_TOT, na.rm = TRUE),
    #sum_le_euro_tot  = sum(LE_EURO_TOT, na.rm = TRUE),
    n_vessels        = n_distinct(VE_ID, na.rm = TRUE),
    vessel_ids       = ifelse (
      n_distinct(VE_ID) < 3,
      paste(unique(VE_ID), collapse = ";"),
      'not_required'
    )
  ) %>% relocate(n_vessels, vessel_ids, .before = Csquare) %>%
  mutate(AverageGearWidth = NA %>% as.numeric()) %>%
  as.data.frame()

write.csv(table1.2, paste0(outPath, year, "_table1_2.csv"))
```

Data product 3 is grouped by Month, 0.05 degree Csquare and gear type.

```{r data product 3, include=TRUE, eval=FALSE}
### Month, 0.05 Csquare, gear

table1.3 = 
  table1 %>%
  group_by(VE_COU,Month,Csquare,LE_GEAR) %>%
  summarise(
    mean_si_sp       = mean(SI_SP),
    sum_intv         = sum(INTV, na.rm = TRUE),
    mean_intv        = mean(INTV, na.rm=TRUE),
    mean_ve_len      = mean(VE_LEN, na.rm = TRUE),
    mean_ve_kf       = mean(VE_KW, na.rm = TRUE),
    sum_kwHour       = sum(kwHour, na.rm=TRUE),
    sum_le_kg_tot    = sum(LE_KG_TOT, na.rm = TRUE),
    #sum_le_euro_tot  = sum(LE_EURO_TOT, na.rm = TRUE),
    n_vessels        = n_distinct(VE_ID, na.rm = TRUE),
    vessel_ids       = ifelse (
      n_distinct(VE_ID) < 3,
      paste(unique(VE_ID), collapse = ";"),
      'not_required'
    )
  ) %>% relocate(n_vessels, vessel_ids, .before = Csquare) %>%
  mutate(AverageGearWidth = NA %>% as.numeric()) %>%
  as.data.frame()

write.csv(table1.3, paste0(outPath, year, "_table1_3.csv"))
```

Data product 4 is grouped by Quarter, 0.01 degree Csquare, gear type and length category.

```{r data product 4, include=TRUE, eval=FALSE}
### Quarter, 0.01 Csquare, gear, length category

table1.4 = 
  table1 %>%
  group_by(VE_COU,QUARTER,Csquare_01,LE_GEAR,LENGTHCAT) %>%
  summarise(
    mean_si_sp       = mean(SI_SP),
    sum_intv         = sum(INTV, na.rm = TRUE),
    mean_intv        = mean(INTV, na.rm=TRUE),
    mean_ve_len      = mean(VE_LEN, na.rm = TRUE),
    mean_ve_kf       = mean(VE_KW, na.rm = TRUE),
    sum_kwHour       = sum(kwHour, na.rm=TRUE),
    sum_le_kg_tot    = sum(LE_KG_TOT, na.rm = TRUE),
    #sum_le_euro_tot  = sum(LE_EURO_TOT, na.rm = TRUE),
    n_vessels        = n_distinct(VE_ID, na.rm = TRUE),
    vessel_ids       = ifelse (
      n_distinct(VE_ID) < 3,
      paste(unique(VE_ID), collapse = ";"),
      'not_required'
    )
  ) %>% relocate(n_vessels, vessel_ids, .before = Csquare_01) %>%
  #mutate(AverageGearWidth = NA %>% as.numeric()) %>%
  as.data.frame()

write.csv(table1.4, paste0(outPath, year, "_table1_4.csv"))
```

Data product 5 is grouped by Year, 0.01 degree Csquare, gear type and length category.

```{r data product 5, include=TRUE, eval=FALSE}
### Year, 0.01 Csquare, gear, length category

table1.5 = 
  table1 %>%
  group_by(VE_COU,Year,Csquare_01,LE_GEAR,LENGTHCAT) %>%
  summarise(
    mean_si_sp       = mean(SI_SP),
    sum_intv         = sum(INTV, na.rm = TRUE),
    mean_intv        = mean(INTV, na.rm=TRUE),
    mean_ve_len      = mean(VE_LEN, na.rm = TRUE),
    mean_ve_kf       = mean(VE_KW, na.rm = TRUE),
    sum_kwHour       = sum(kwHour, na.rm=TRUE),
    sum_le_kg_tot    = sum(LE_KG_TOT, na.rm = TRUE),
    #sum_le_euro_tot  = sum(LE_EURO_TOT, na.rm = TRUE),
    n_vessels        = n_distinct(VE_ID, na.rm = TRUE),
    vessel_ids       = ifelse (
      n_distinct(VE_ID) < 3,
      paste(unique(VE_ID), collapse = ";"),
      'not_required'
    )
  ) %>% relocate(n_vessels, vessel_ids, .before = Csquare_01) %>%
  #mutate(AverageGearWidth = NA %>% as.numeric()) %>%
  as.data.frame()

write.csv(table1.5, paste0(outPath, year, "_table1_5.csv"))
```

Data product 6 is grouped by Month, 0.01 degree Csquare, gear type and length category.

```{r data product 6, include=TRUE, eval=FALSE}
### Month, 0.01 Csquare, gear, length category

table1.6 = 
  table1 %>%
  group_by(VE_COU,Month,Csquare_01,LE_GEAR,LENGTHCAT) %>%
  summarise(
    mean_si_sp       = mean(SI_SP),
    sum_intv         = sum(INTV, na.rm = TRUE),
    mean_intv        = mean(INTV, na.rm=TRUE),
    mean_ve_len      = mean(VE_LEN, na.rm = TRUE),
    mean_ve_kf       = mean(VE_KW, na.rm = TRUE),
    sum_kwHour       = sum(kwHour, na.rm=TRUE),
    sum_le_kg_tot    = sum(LE_KG_TOT, na.rm = TRUE),
    #sum_le_euro_tot  = sum(LE_EURO_TOT, na.rm = TRUE),
    n_vessels        = n_distinct(VE_ID, na.rm = TRUE),
    vessel_ids       = ifelse (
      n_distinct(VE_ID) < 3,
      paste(unique(VE_ID), collapse = ";"),
      'not_required'
    )
  ) %>% relocate(n_vessels, vessel_ids, .before = Csquare_01) %>%
  #mutate(AverageGearWidth = NA %>% as.numeric()) %>%
  as.data.frame()

write.csv(table1.6, paste0(outPath, year, "_table1_6.csv"))
```

###Data Product Visialisation

Find the bounding box of the original table (table1)

```{r lon/lat, include=TRUE, eval=FALSE}
## bounding box area limits definition

# find the min and max limits of the table's data
lats = c ( min(table1$SI_LATI), max(table1$SI_LATI) ) 
lons = c ( min(table1$SI_LONG), max(table1$SI_LONG) ) 

lats
lons
```

Overwrite this information by expanding it, to give a better field of view.

```{r aoi, include=TRUE, eval=FALSE}
## Welsh boundaries
lats = c ( 51, 53.5) 
lons = c ( -6, -2 ) 
```

Create a bounding box using these longitudes and latitudes.

```{r bounding box, include=TRUE, eval=FALSE}
coord_bbox = data.frame(lons, lats)
bbox_aoi =  coord_bbox %>% 
  st_as_sf(coords = c("lons","lats"), crs = 4326) %>%
  st_bbox() %>% st_as_sfc() %>% 
  st_sf( id  = 1, label = 'bbox' ) %>%
  st_set_crs(4326)
```

#### create the grids at different resolutions

```{r 0.05 resolution, include=TRUE, eval=FALSE}
cell_resolution_005 = 0.05 
csquare_resolution_005 = 0.05

##spatial grid 0_05 

grid_0_05  = st_make_grid( bbox_aoi,  cellsize = cell_resolution_005, square = TRUE, offset = c(min(lons),min(lats)))#%>% as(Class="Spatial")
 
grid_0_05_centroid  = st_sf (grid_0_05) %>%  st_centroid (.) %>% st_coordinates() 
grid_0_05csquares =  CSquare ( grid_0_05_centroid[,"X"], grid_0_05_centroid[,"Y"],  csquare_resolution_005 )

csquares_0_05_split_centroids = st_bind_cols( grid_0_05, 
                                         grid_res = cell_resolution_005, 
                                         csquare = grid_0_05csquares, 
                                         lon = round ( grid_0_05_centroid[,"X"],5), 
                                         lat = round ( grid_0_05_centroid[,"Y"],5))
```


```{r 0.01 resolution, include=TRUE, eval=FALSE}
cell_resolution_001 = 0.01 
csquare_resolution_001 = 0.01

grid_0_01  = st_make_grid( bbox_aoi,  cellsize = cell_resolution_001, square = TRUE, offset = c(min(lons),min(lats)))#%>% as(Class="Spatial")

grid_0_01_centroid  = st_sf (grid_0_01) %>%  st_centroid (.) %>% st_coordinates() 
grid_0_01csquares =  CSquare ( grid_0_01_centroid[,"X"], grid_0_01_centroid[,"Y"],  csquare_resolution_001 )

csquares_0_01_split_centroids = st_bind_cols( grid_0_01, 
                                              grid_res = cell_resolution_001, 
                                              csquare = grid_0_01csquares, 
                                              lon = round ( grid_0_01_centroid[,"X"],5), 
                                              lat = round ( grid_0_01_centroid[,"Y"],5))
```

Read in the world shapefile. If you have the naturalearth package for R, you can call the world shapefile using the hashed out code. Otherwise, the read_sf line will read it in from a local source. Additionally, the crs of the shapefile is changed to ESPG: 4326.

```{r world geometry, include=TRUE, eval=FALSE}
# read in the world shapefile, or use the ne_countries function from the natural earth package if available
world = read_sf("C:\\folder\\where\\shapefile\\is\\stored\\europe_coastline_folder", "europe_coastline") %>% st_transform(., crs = 4326)
#world <- ne_countries(scale = "large", returnclass = "sf")
```

Add the spatial grid geometry to the fisheries data products tables.

```{r assign spatial grid, include=TRUE, eval=FALSE}
table1.1_geom = csquares_0_05_split_centroids %>% inner_join(table1.1, by = c("csquare" = "Csquare"))
table1.2_geom = csquares_0_05_split_centroids %>% inner_join(table1.2, by = c("csquare" = "Csquare"))
table1.3_geom = csquares_0_05_split_centroids %>% inner_join(table1.3, by = c("csquare" = "Csquare"))
table1.4_geom = csquares_0_01_split_centroids %>% inner_join(table1.4, by = c("csquare" = "Csquare_01"))
table1.5_geom = csquares_0_01_split_centroids %>% inner_join(table1.5, by = c("csquare" = "Csquare_01"))
table1.6_geom = csquares_0_01_split_centroids %>% inner_join(table1.6, by = c("csquare" = "Csquare_01"))
```

Save the _geom outputs as geojson files.

```{r save as geojson, include=TRUE, eval=FALSE}
st_write(table1.1_geom, paste0(gjPath, "table1_1_geom.geojson"), layer = "table1_1_geom.geojson") ## save as geojson 
st_write(table1.2_geom, paste0(gjPath, "table1_2_geom.geojson"), layer = "table1_2_geom.geojson")
st_write(table1.3_geom, paste0(gjPath, "table1_3_geom.geojson"), layer = "table1_3_geom.geojson")
st_write(table1.4_geom, paste0(gjPath, "table1_4_geom.geojson"), layer = "table1_4_geom.geojson")
st_write(table1.5_geom, paste0(gjPath, "table1_5_geom.geojson"), layer = "table1_5_geom.geojson")
st_write(table1.6_geom, paste0(gjPath, "table1_6_geom.geojson"), layer = "table1_6_geom.geojson")
```

### Create Data Products

Using the function written at the start of the script, data products will now be made. res1 will be the name of the data product made using table1.1. The required elements are selected here in the inverted commas.

```{r create data product 1, include=TRUE, eval=FALSE}
res1 = plot_function(data_plot = table1.1_geom, col_plot = 'sum_intv', col_facet1 = "LE_GEAR", col_facet2 = "QUARTER")
```

Save the data product to the plots folder created at the beginning. The width and height are specified so that they should fill the whole screen, with a resolution of 300 dpi for a balance of good detail and short enough processing time.

```{r save data product 1, include=TRUE, eval=FALSE}
ggsave(plot = res1, filename = paste0("table1_1_plot.png"), path = plotsPath, width = 16, height = 9, dpi = 300)
```

Repeat these two steps for the rest of the data products.

```{r create data products, include=TRUE, eval=FALSE}
res2 = plot_function(data_plot = table1.2_geom, col_plot = 'sum_intv', col_facet1 = "LE_GEAR", col_facet2 = "Year")
ggsave(plot = res2, filename = paste0("table1_2_plot.png"), path = plotsPath, width = 16, height = 9, dpi = 300)

res3 = plot_function(data_plot = table1.3_geom, col_plot = 'sum_intv', col_facet1 = "LE_GEAR", col_facet2 = "Month")
ggsave(plot = res3, filename = paste0("table1_3_plot.png"), path = plotsPath, width = 16, height = 9, dpi = 300)

res4 = plot_function(data_plot = table1.4_geom, col_plot = 'sum_intv', col_facet1 = "LE_GEAR", col_facet2 = "QUARTER")
ggsave(plot = res4, filename = paste0("table1_4_plot.png"), path = plotsPath, width = 16, height = 9, dpi = 300)

res5 = plot_function(data_plot = table1.5_geom, col_plot = 'sum_intv', col_facet1 = "LE_GEAR", col_facet2 = "Year")
ggsave(plot = res5, filename = paste0("table1_5_plot.png"), path = plotsPath, width = 16, height = 9, dpi = 300)

res6 = plot_function(data_plot = table1.6_geom, col_plot = 'sum_intv', col_facet1 = "LE_GEAR", col_facet2 = "Month")
ggsave(plot = res6, filename = paste0("table1_6_plot.png"), path = plotsPath, width = 16, height = 9, dpi = 300)
```
