---
title: "1_Load_TACSAT_and_EFLALO_in_R_environment"
author: "Roi & Mike"
date: "2023-09-14"
output: html_document
---

```{r setup, include=TRUE, eval=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## Load required libraries

library(lubridate)  ## this package helps with manage date fields 
library(dplyr)      ## R Package to use pipelines (%>%) analysis language 
library(sf)         ## R Package for spatial analysis in R ( Simple Features )
library(ggplot2)    ## R Package for plotting and graphs
```

## Load eflalo data blocks and merge them into one dataframe

Check the working directory, where all paths are based from, by default.Read in the eflalo_ft data block. Header = F tells R whether or not to expect the first row to be the headers for each column. If this were set to T for TRUE, it would assume the first row of data was the column names. The sep = part of the function is used as the data will be presented in a csv format, so the values in each row are separated by commas, which is chosen here.

```{r set wd, include=TRUE, eval=FALSE}
  ##check WD is in the desired data folder location
  getwd() 
  
  ## otherwise change to desired data folder location
  setwd('./../../data')
```

Next, select the analysis type. Analysis is separated into trips in Welsh waters, and Welsh vessels. Welsh waters analysis will investigate only the data which intersects with the Welsh zone, regardless of the vessel's nationality. Welsh waters analysis will set the file path leading to the data containing only welsh vessels. If statements have a requirement that, if met, will progress the code in a certain way. The else portion of the statement dictates what will happen if the parameter is not met or if a different parameter is met.

```{r select analysis, include=TRUE, eval=FALSE}
 ## SELECT THE ANALYSIS OPTION:
  
  analysis_type = 'welsh_fleet' ## Options: ( 'welsh_fleet' , 'welsh_waters'  ) 
  
  if ( analysis_type == 'welsh_fleet')  { 

    ###  1. WELSH FLEET ACTIVITY ANALYSIS 
    
    data_folder_t3 = file.path( paste0 ( getwd() , '/welsh_fleet_data/t3' )  )
    data_folder_geofish = file.path( paste0 ( getwd() , '/welsh_fleet_data/geofish' ) ) 

  } else if  (analysis_type == 'welsh_waters' ) { 


  ## 2. WELSH WATERS ACTIVITY ANALYSIS 
  
  
  data_folder_t3 = file.path( paste0 ( getwd() , '/welsh_waters_data/t3' )  )
  data_folder_geofish = file.path( paste0 ( getwd() , '/welsh_waters_data/geofish' ))
  
  } 
```

Read in the eflalo_ft data block. Header = T tells R whether or not to expect the first row to be the headers for each column. If this were set to F for FALSE, it would assume the first row of data was data, not the column names. The sep = part of the function is used as the data will be presented in a csv format, so the values in each row are separated by commas, which is chosen here. Finally, set column names to upper case to meet formatting requirements.

```{r load eflalo_ft, include=TRUE, eval=FALSE}
eflalo_ft_t3 = read.csv(file = paste0(data_folder_t3, '\\eflalo_ft.csv') , header = T, sep = ','  , fileEncoding = 'UTF-8-BOM')
eflalo_ft_gf = read.csv(file = paste0(data_folder_geofish, '\\eflalo_ft.csv') , header = T, sep = ','  , fileEncoding = 'UTF-8-BOM')
names(eflalo_ft_gf) = toupper( names(eflalo_ft_gf ) )  ## Column names are lower case in geofish , needs to be changed to upper case 
```

Data is coming from two sources: GeoFISH and Tier3. The column which identifies the vessel's country is named and formatted differently between the two. Here, the Tier3 format is converted to match the GeoFISH format, using three letter codes to identify vessel country of origin.

```{r convert ve_cou, include=TRUE, eval=FALSE}
## Convert the field VE_COU in a standard country acronym 
countries_id = data.frame( ve_fa =   c('Wales', 'England', 'Isle of Man', 'NULL', 'Scotland', 'North Ireland') , ve_cou_gb = c( 'GBW', 'GBE', 'GBI','NULL', 'GBS', 'GBN' ) ) 
eflalo_ft_t3 = eflalo_ft_t3 %>% left_join(countries_id, by =  c ( 'VE_FA' = 've_fa')) %>% mutate(VE_COU = ve_cou_gb) %>% select ( - ve_cou_gb)
```

Filter the data to only include vessels over 10m from GeoFISH and under 10m from Tier3, for the most accurate information.

```{r filter by vessel length, include=TRUE, eval=FALSE}
## Filter GeoFISH data for Over 10 meter vessesl. Logbook info from <10 m in GeoFISH it comes from Sales Notes and is less reliable.

eflalo_ft_gf = eflalo_ft_gf %>% filter ( VE_LEN >= 10 )
eflalo_ft_t3 = eflalo_ft_t3 %>% filter ( as.numeric( VE_LEN)  < 10 )
```

Explore the data with these summarisation commands.

```{r summarise eflalo ft, include=TRUE, eval=FALSE}
  head (eflalo_ft_t3)
  str(eflalo_ft_t3)
  dim(eflalo_ft_t3)
  summary(eflalo_ft_t3)
```

Now load the eflalo le data and inspect it. The class of the ICES rectangle column is specified as character, to avoid it being numeric and converting to a number.

```{r load eflalo le, include=TRUE, eval=FALSE}
# load eflalo_le from both sources ----
  eflalo_le_t3 = read.csv(file = paste0(data_folder_t3, '\\eflalo_le.csv' ) , header = T, sep = ',',
                          fileEncoding = 'UTF-8-BOM', colClasses = c ( rep(NA, 10) , "character"))  

  eflalo_le_gf = read.csv(file = paste0(data_folder_geofish, '\\eflalo_le.csv' ) , header = T, sep = ',',
                          fileEncoding = 'UTF-8-BOM' , colClasses = c ( rep(NA, 10) , "character"))
  
  names(eflalo_le_gf) = toupper( names(eflalo_le_gf ) )  ## Column names are lower case in geofish , needs to be changed to upper case 

  head (eflalo_le_t3)
  str (eflalo_le_t3)
```

Load eflalo_spe data.

```{r load eflalo spe, include=TRUE, eval=FALSE}
# load eflalo_spe from both sources ----
eflalo_spe_t3 = read.csv(file = paste0(data_folder_t3, '\\eflalo_spe.csv' ), header = T, sep = ','  , fileEncoding = 'UTF-8-BOM')
eflalo_spe_gf = read.csv(file = paste0(data_folder_geofish, '\\eflalo_spe.csv' ) , header = T, sep = ','  , fileEncoding = 'UTF-8-BOM')
names(eflalo_spe_gf) = toupper( names(eflalo_spe_gf ) )  ## Column names are lower case in geofish , needs to be changed to upper case 
```

Join the blocks together to make the complete EFLALO dataset. An inner join retains all data rows that have a match in the other columns. Check the results with head() and str().

```{r join eflalo, include=TRUE, eval=FALSE}
eflalo_t3  =    eflalo_ft_t3 %>%
                inner_join (eflalo_le_t3 , by =  c("FT_REF" = "EFLALO_FT_FT_REF"))%>%
                inner_join(eflalo_spe_t3, by = c("LE_ID" = "EFLALO_LE_LE_ID"   ))


eflalo_t3 = eflalo_t3 %>% mutate ( LE_VALUE = -9999, FLEET_SEG = analysis_type, SOURCE = 't3') %>% select  ( - VE_FA)

eflalo_gf =     eflalo_ft_gf %>%
                inner_join (eflalo_le_gf , by =  c("FT_REF" = "EFLALO_FT_FT_REF"))%>%
                inner_join(eflalo_spe_gf, by = c("LE_ID" = "EFLALO_LE_LE_ID"   ))

eflalo_gf = eflalo_gf %>% rename ( LE_VALUE = LE_EURO )  %>%  mutate (   FLEET_SEG = analysis_type, SOURCE = 'geofish') 


eflalo = rbind(eflalo_t3, eflalo_gf)
```

Some field require some formatting. the ymd() function orientates the selected columns with dates, into a year-month-day format. Columns with time values associated can be formatted using ymd_hms.

The vessel length field (VE_LEN) is required to be in numerical format; as.numeric sorts that out.

Finally, the year and month columns are created by pulling the information from the previously formatted columns.

```{r eflalo column formatting, include=TRUE, eval=FALSE}
eflalo$FT_DDAT =  ymd(eflalo$FT_DDAT)  ## ymd lubridate function to CAST date into Year Mond Day date format
eflalo$FT_LDAT =  ymd(eflalo$FT_LDAT) 

eflalo = eflalo %>% mutate (LE_CDAT  = substr (eflalo$LE_CDAT, 1, 10) )  ## To convert Log Event date just in Year- month - day. Time is not provided an is by default 00:00:00 
eflalo$LE_CDAT = ymd(eflalo$LE_CDAT) 

eflalo$FT_DDATIM = ymd_hms(eflalo$FT_DDATIM) 
eflalo$FT_LDATIM = ymd_hms(eflalo$FT_LDATIM) 

eflalo$VE_LEN = as.numeric(eflalo$VE_LEN)
eflalo$VE_KW = as.numeric(eflalo$VE_KW)
eflalo$VE_TON = as.numeric(eflalo$VE_TON)

eflalo$EFLALO_FT_FT_REF  = as.numeric(eflalo$EFLALO_FT_FT_REF)
eflalo$LE_VALUE = as.numeric(eflalo$LE_VALUE )

eflalo$Year = year(eflalo$FT_DDATIM )
eflalo$Month = month(eflalo$FT_LDATIM)
```

## Load Tacsat data

Load data from GeoFISH and Tier3.

```{r load tacsat, include=TRUE, eval=FALSE}
tacsat_t3 = read.csv(file = paste0(data_folder_t3, '\\tacsat.csv' ) , header = T, sep = ','  , fileEncoding = 'UTF-8-BOM')
tacsat_t3 = tacsat_t3 %>% mutate ( FLEET_SEG = analysis_type, SOURCE = 't3') 


tacsat_gf = read.csv(file = paste0(data_folder_geofish, '\\tacsat.csv' ) , header = T, sep = ','  , fileEncoding = 'UTF-8-BOM')
names(tacsat_gf) = toupper( names(tacsat_gf ) )  ## Column names are lower case in geofish , needs to be changed to upper case 
tacsat_gf = tacsat_gf %>% mutate ( FLEET_SEG = analysis_type, SOURCE = 'geofish') 
```

Bind the tacsat datasets together and investigate them.

```{r bind tacsat, include=TRUE, eval=FALSE}
tacsat = rbind(tacsat_t3, tacsat_gf)
                     
 
dim(tacsat)
head (tacsat)
str(tacsat) ##List of fields and field type
```

Format dates, times and numerical columns for consistency purposes. These will be in the same format as the EFLALO data. Also filter to see how many NAs are present in the speed column (SI_SP). This does not change the data as it is not applying the command to the object. It simply prints the number of rows containing NAs. If there are a large number, this can be investigated.

```{r format tacsat columns, include=TRUE, eval=FALSE}
tacsat$SI_DATE  =  ymd( tacsat$SI_DATE   )  ### reformatting the data in required format . Change to dmy if your system date format is different
tacsat$SI_DATIM  = ymd_hms(tacsat$SI_DATIM  ) 
tacsat$SI_SP = as.numeric(tacsat$SI_SP)
tacsat$SI_HE = as.numeric(tacsat$SI_HE)
```

If the analysis being conducted is the Welsh waters analysis, the following code will extract the data intersecting with the Welsh zone. If it is Welsh vessels analysis being conducted, this code will just be passed over. The code block reads the welsh plan area geojson, intersecting the vms data with the boundary. Those inside the boundary are kept.

```{r welsh waters analysis, include=TRUE, eval=FALSE}
if ( analysis_type == 'welsh_waters' ) { 
  
  welsh_marine_area = st_read ( dsn = '.\\spatial_layers\\wales_plan_area.geojson' )
  welsh_marine_area_geom = st_make_valid(st_union(welsh_marine_area  ))
  
  
  tacsat_geom = tacsat %>% st_as_sf( ., coords = c("SI_LONG" ,"SI_LATI"), crs = 4326 , remove = FALSE)
  tacsat_geom_ww =  tacsat_geom  %>%  filter(  st_intersects( .,  welsh_marine_area_geom , sparse = FALSE)  )
  
  
  trips_in_welsh_waters = tacsat_geom_ww %>% st_drop_geometry() %>%  distinct(SI_FT ) %>% pull()
  
  eflalo = eflalo %>% filter ( FT_REF  %in%  trips_in_welsh_waters) 
  tacsat = tacsat  %>% filter ( SI_FT  %in% trips_in_welsh_waters ) 
  

  ## To visualize in a GIS software save the TACSAT  as point geometry 
  
  

  dir.create(".\\workflow_outputs\\spatial")

  st_write( tacsat_geom_ww, dsn = ".\\workflow_outputs\\spatial\\tacsat_welsh_waters.geojson", layer = "tacsat_welsh_waters.geojson")
  
  
}
```

Save the outputs to the workflow_outputs folder.

```{r save outputs, include=TRUE, eval=FALSE}
save(eflalo, file = ".\\workflow_outputs\\eflalo.RData")
save(tacsat, file = ".\\workflow_outputs\\tacsat.RData")
```
