---
title: "0_eflalo_tacsat_preprocessing"
author: "Roi & Mike"
date: "2023-09-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## Load required libraries

library(lubridate)  ## this package helps with manage date fields 
library(dplyr)      ## R Package to use pipelines (%>%) analysis language 
library(sf)         ## R Package for spatial analysis in R ( Simple Features )
library(ggplot2)    ## R Package for plotting and graphs
```

Set the working directory if it is not already set to the data folder.

```{r set wd, include=FALSE}
setwd('./../../data')
```

Load the data created via by the previouis workflow.

```{r load data, include=FALSE}
load(".\\workflow_outputs\\eflalo.RData")
load(".\\workflow_outputs\\tacsat.RData")
```

Select fleet segment. Similarly to the previous if statement, this section selects which segment of the dataset is being analised. The options are Over 12m vessels, Under 12m vessels and all vessels.

```{r select fleet segment, include=FALSE}
fleet_segment = 'under12' ## Options: ( 'over12', 'under12', 'all' )
 

if ( fleet_segment == 'over12')  { 
  
  ###  1. Over 12m vessels analysis
  
  eflalo_fs = eflalo %>% filter(VE_LEN >= 12) %>% mutate  ( FLEET_SEG = paste0 ( FLEET_SEG, '_', fleet_segment))
  tacsat_fs = tacsat %>% filter(VE_REF %in% eflalo_fs$VE_REF)
  #tacsat = tacsat %>% filter(SI_FT %in% eflalo$FT_REF)
  
  
} else if  (fleet_segment == 'under12' ) { 
  
  
  ### 2. Under 12m vessels analysis 
  
  eflalo_fs = eflalo %>% filter(VE_LEN < 12) %>% mutate  ( FLEET_SEG = paste0 ( FLEET_SEG, '_', fleet_segment))
  tacsat_fs = tacsat %>% filter(VE_REF %in% eflalo_fs$VE_REF)
  
} else if  (fleet_segment == 'all' ) { 
  
  
  ### 3. All vessels 
  
  eflalo_fs = eflalo 
  tacsat_fs = tacsat  
  
} 
```

## Explore the Eflalo data

Filter the columns to find all of the entries listed only once each, regardless of how many times they occur in the column. Secondly, filter the eflalo_fs. dataset to vessels where the VE_COU column is listed as 'GBE'. These entries are assigned to eflalo_fs, which comes before the =. Using the %>% pipeline from the dplyr package, the object which will have the actions done to it is listed first after the =. Following this, the filter function is applied to eflalo_fs. Usually the format would be filter(eflalo_fs., VE_COU == 'GBE'), but as eflalo_fs. has been specified before the pipe, it is not included in the brackets as R already knows to apply the function to this object. Sometimes, programmers will put a . in place on the object, as a way of signifying the object again, e.g. eflalo_fs. %>% filter(., VE_COU == 'GBE'). But this is not technically necessary.

```{r explore vessels, include=FALSE}
## Explore the stats from EFLALO to get fleet statistics and fleet characterisation


  ## Explore vessels nationality in the EFLALO records 

  eflalo_fs %>% distinct(VE_COU)
  eflalo_fs %>% distinct(FT_DCOU)
  eflalo_fs %>% distinct(VE_REF)
  
  ## Filter (if necessary) the vessels from nationality required using the Vessel Country field (VE_COU)
  # The country code can be changed to your requirements, this is not editing the data, just exploring it
  eflalo_fs %>% filter(VE_COU == 'GBE')
  eflalo_fs %>% group_by(VE_COU, SOURCE )%>% tally ( ) %>% arrange(VE_COU)
  
  
  ## Check the size change in dimensions of the dataset you chose
  
  eflalo_fs %>% distinct(FT_REF) %>% dim()
```

Here, multiple pipes are used, meaning multiple functions are applied to the object at the beginning: eflalo_fs. However, the code does not follow an = sign, meaning the outcome is not being applied to a new object, so it will just be printed to the console.

Here, vessels 10m or greater in length, tht came from GeoFISH are selected, to see the distribution of vessels between sources. Summary statistics are provided for the vessel lengths and graphs are plotted to display these.

```{r check vessel lengths, include=FALSE}
  ## Check vessel length categories
  
  eflalo_fs %>% filter( VE_LEN >= 10  & SOURCE == 'geofish')
  
  eflalo_fs %>% select(VE_LEN)%>%mutate( VE_LEN = as.numeric(VE_LEN)) %>% summary()
  eflalo_fs %>% mutate ( VE_LEN =  floor ( VE_LEN ) )  %>% distinct(VE_LEN, SOURCE ) %>% group_by(VE_LEN , SOURCE )%>% tally ()
  
  eflalo_fs  %>% distinct(VE_REF, VE_LEN) %>% mutate( VE_LEN = floor(VE_LEN  )) %>% group_by(VE_LEN) %>% tally() %>%
    ggplot(., aes(VE_LEN, n)) + geom_bar(stat = "identity") +scale_x_continuous( breaks = seq(1:32) )
  
  eflalo_fs %>% ggplot( . , aes( VE_LEN)) + geom_histogram()
```

Create vessel length categories based on the fleet segment. The chosen segment dictates the category distribution. This is done via an ifelse statement, as above. VE_LEN_CAT does not yet exist, so it is created here, using what follows the =. Categories are created using breaks, as the categories will reside in between each break value. E.g. the first category is -Inf to 4.5, meaning anything below 4.5 is assigned here. Then anything between 4.5 and 5 is assigned to the next category, and so on. Finally, the labels are assigned in the same order, so it is important they are correctly located in the list. 

```{r vessel length categories, include=FALSE}
  ## Create the VESSEL LENGTH CAGTEGORY to explore data by vessel length classes
  
    if ( fleet_segment == 'over12')  {  
    
      eflalo_fs$VE_LEN_CAT = cut(eflalo_fs$VE_LEN , include.lowest = T,
                                 breaks=c(-Inf, 12, 15, 18, 24, 40,  Inf), 
                                 labels=c("<12m",  "12-15m","15-18m", "18-24m", "24-40m" ,"=>40m"))
      
      
      
    
    
    } else if  (fleet_segment == 'under12' ) { 
    
    
      eflalo_fs$VE_LEN_CAT = cut(eflalo_fs$VE_LEN , include.lowest = T,
                                 breaks=c(-Inf, 4.5, 4.999 , 6.999, 8.999, 11.999 , Inf), 
                                 labels=c("<4.5m","4.5-5m", "5-7m", "7-9m", "9-12m", "=>12m"))
    
    } else if ( fleet_segment == 'all') { 
      
      
      eflalo_fs$VE_LEN_CAT = cut(eflalo_fs$VE_LEN , include.lowest = T,
                                 breaks=c(-Inf, 6, 8, 10, 12, 15, 18, 24, 40,  Inf), 
                                 labels=c("<6m","6-8m", "8-10m", "10-12m", "12-15m","15-18m", "18-24m", "24-40m" ,"=>40m"))
      
    }

  eflalo_fs %>% select ( FT_REF , VE_REF, VE_LEN, VE_LEN_CAT)
```

Create the trip days field by calculating the difference between departure and landing time.

```{r , include=FALSE}
  ## Create the field "trip_days" with duration of each trips as   number of days  

  eflalo_fs  = eflalo_fs %>% mutate ( trip_days = as.numeric(difftime(eflalo_fs$FT_LDATIM, eflalo_fs$FT_DDATIM), units = "days") ) 
  
  head(eflalo_fs)
  str(eflalo_fs)
```

## Plotting 

Plot the categories/columns to understand the fleet composition.

Plot 1 is a bar graph which shows the number of trips made per vessel category.

```{r plot 1, include=FALSE}
  ## Plot the categories to understand your fleet composition
  
    ## PLOT 1: Number of trips by vessel length category 
    
    ggplot(data = eflalo_fs %>% distinct(FT_REF, VE_LEN_CAT), aes(VE_LEN_CAT)) +
    geom_bar()
```

Plot 2 shows the number of trips made per gear type.

```{r plot 2.1, include=FALSE}
    ## PLOT 2.1: Number of trips by gear category 
    
    ggplot(data = eflalo_fs%>%distinct(FT_REF, LE_GEAR), aes(LE_GEAR)) +
    geom_bar()
```

Another way of doing this is using facet wrap. This will create individual plots for each element of the chosen category, which in this case is VE_LEN_CAT.

```{r plot 2.2, include=FALSE}
    ## PLOT 2.2: Number of trips by gear category, using facet wrap for each length category
    
    ggplot(data = eflalo_fs%>%distinct(FT_REF,VE_LEN_CAT, LE_GEAR), aes(LE_GEAR)  ) +
      geom_bar() + 
      facet_wrap( ~ VE_LEN_CAT)
```

```{r plot 2.3, include=FALSE}
    ## Now for each gear type
    
    ggplot(data = eflalo_fs%>%distinct(FT_REF,VE_LEN_CAT, LE_GEAR) , aes(VE_LEN_CAT)  ) +
      geom_bar() + 
      facet_wrap( ~ LE_GEAR)
```

The object res1 is created by summing the total weight of each species caught and each gear type used. These elements are chosen using the group_by function, which tells R to put entries with the same values in these columns together, summing the contents of their LE_KG columns and assigning the output to a new column: le_kg_total. Finally, the rows are sorted in descending order by total weight, but separated by gear type, so the same gear types can be compared easily. 

```{r plot 3, include=FALSE}
    ## PLOT 3: Main species captured
    
      ## Rank the main species captured by gear type
    
    res1 = eflalo_fs %>%
    group_by(  LE_GEAR, LE_SPE) %>%
    summarise(le_kg_total = sum( LE_KG) )  %>%
    mutate ( rank = dense_rank( desc(le_kg_total) )) %>%
    arrange(LE_GEAR, desc(le_kg_total) ) %>%
    filter(rank <= 5) %>%
    ungroup() 
    
   
    ggplot( data = res1 , aes( y =le_kg_total, x = LE_SPE)) + 
      geom_bar(stat = "identity") + 
      facet_wrap(~ LE_GEAR, scales = "free" )
```

Plot 4 filters for trips which are claiming extremely long duration. These are plotted via histogram. Secondly, trips of less than 50 days are plotted.

```{r plot 4, include=FALSE}
    ## PLOT 4: Fishing trip duration and fish duration by vessel length. 
    
      #Q1: Do we expect longer trips from larger vessels?
      #Q2: Is there any trips duration identified as outliers? 
    
    ggplot(data = eflalo_fs %>% filter( trip_days < 500 ) ) +
    geom_histogram( aes( trip_days ) )
    
    ggplot(data = eflalo_fs %>% filter(trip_days < 50) , aes ( x = trip_days, y = VE_LEN , color = SOURCE )) + 
    geom_point(  )
```

Save outputs.

```{r , include=FALSE}
    write.csv( x = res1, file =  ".\\workflow_outputs\\species_kg_ranked_by_gear.csv", row.names=FALSE)
    
    ## Save the intermediate EFLALO and TACSAT datasets
    
    save(eflalo_fs, file = ".\\workflow_outputs\\eflalo_fs.RData")
    save(tacsat_fs, file = ".\\workflow_outputs\\tacsat_fs.RData")
```
